---
title: "1 - Preparing and Exploring Variables in ARG data"
author: "Morgane Gicquel"
date: "2025-08-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare workspace}
source("scripts/source_packages.R")
source("scripts/source_themes_mg.R") # ggplot2 themes

wrkdir <- getwd()
output_wd <- paste0(wrkdir, "/output")
geoproc_wd <- paste0(wrkdir, "/data/processed/geo")
```

# Loading data files

## Taxonomic profile results
PS.mOTUS.rds  
  otu_table: Matrix with raw read counts per mOTU per sample (Mus musculus)  
  tax_table: Matrix with full taxonomy by mOTU  
  sam_table: Unified metadata for all samples  
  
## ARG profile results
PS.ARG.raw.rds  
  otu_table: Matrix with raw read counts per ARG per sample (Mus musculus)  
  tax_table: Matrix with full ARG annotation  
  sam_table: Unified metadata for all samples  
  
PS.ARG.FPKM.rds  
  otu_table: Matrix with ARG counts normalized to fragments per kilobase per million mapped reads per sample (Mus musculus)  
  tax_table: Matrix with full ARG annotation  
  sam_table: Unified metadata for all samples  

Best to use FPKM as corrected for genes lengths.


```{r load data}

# Filtered data

PS.mOTUS.filt <- readRDS("./data/raw/PS.mOTUS.filtered.rds")

PS.ARG.raw.filt <- readRDS("./data/raw/PS.ARG.filtered.rds")

PS.ARG.FPKM.filt <- readRDS("./data/raw/PS.ARG.FPKM.filtered.rds")

```

# Bacteria

### Cleaning up NA in Taxa table
```{r}

PS.mOTUS.filt

# let's clean up genus column in the tax table

tax <- as.data.frame(tax_table(PS.mOTUS.filt)) %>%
  mutate(
    Phylum = ifelse(str_detect(Phylum, "NA"), NA, Phylum),
    Class = ifelse(str_detect(Class, "NA"), NA, Class),
    Order = ifelse(str_detect(Order, "NA"), NA, Order),
    Family = ifelse(str_detect(Family, "NA"), NA, Family),
    Genus = ifelse(str_detect(Genus, "NA"), NA, Genus),
    Species = ifelse(str_detect(Species, "NA"), NA, Species)) %>%
  mutate(Genus = ifelse(is.na(Genus) | Genus == "Unknown_genus_in_NA", paste0("Unknown_genus_in_", coalesce(Family, Order, Class, Phylum, Kingdom)), Genus)) %>% 
  mutate(Family = ifelse(is.na(Family) | Family == "Unknown_family_in_NA", NA, Family)) %>%
   mutate(Phylum = ifelse(is.na(Phylum), "Unknown_phylum", Phylum),
          Class = ifelse(is.na(Class), "Unknown_class", Class),
          Order = ifelse(is.na(Order), "Unknown_order", Order),
          Genus = ifelse(Genus == "uncultured", paste(Order, Genus, sep = "_"), Genus))

PS.mOTUS.filt@tax_table <- tax_table(as.matrix(tax))

family <- get_taxa_unique(PS.mOTUS.filt, "Family")
family

#View(PS.mOTUS.filt@tax_table)

length(unique(tax_table(PS.mOTUS.filt)[,"Family"]))

```

## Preparing

```{r prepare data}
## N mice trapped per location

aggregateDensity <- unclass(sample_data(PS.mOTUS.filt)) %>%
    ## need to remove class "sample_data" (phyloseq)
    as_tibble() %>%
    ## need to remove NA the same data 
    filter(!is.na(Body_Length)) %>% 
    ## creating Locality variable
    mutate(Locality = paste(round(Latitude,4), " ", round(Longitude,4))) %>%
    ## add total mice trapped per locality and year  
    group_by(Locality, Year) %>%
    ## Number of mice trapped at one locality (combined long/lat at
    ## 4-digit resolution, available in "Locality"). We processed all
    ## mice for each locality. We create the variable per year to
    ## estimate density at one point in time. Hypothesis: as measure
    ## of host population density this indicates increase transmission
    ## between mice. More host-associated transmissible taxa will be
    ## enriched. Potential alternatives not considered now:
    ## collapsing localities at larger scale (more rounding),
    ## persistent density over years for those)
    summarise(Ntrapped = n())  


## filter NAs and select variables

allSData <- unclass(sample_data(PS.mOTUS.filt)) %>%
    ## need to remove class "sample_data" (phyloseq)
    as_tibble() %>%
    ## mice needing to be dropped because of NA values (might be
    ## worthwhile to imput this at some point, "Status" could be most
    filter(!is.na(Body_Length)) %>% 
    ## add here other variable that could need filtering if included
    ## SELECTION FOR NOW:
    ## Latitude and longitude: the basic spatial coordinates, decimal
    ## format, rounded to 4 digits
    ## creating Locality variable
    mutate(Locality = paste(round(Latitude,4), " ", round(Longitude,4)),
           farm_ID = as.integer(factor(Locality))) %>%
    mutate(mouseID = Mouse_ID,
           sex = Sex, 
           lon = Longitude, 
           lat = Latitude, 
           ## year: the year of sampling, hypothesis: closer years are more
           ## similar as a result of community persistence or similarity of
           ## environmental condition in closer years.
           year = as.integer(Year),
           ## BMI: the residuals of body weight on body
           ## length. Hypothesis: mice with worse body condition have
           ## disturbed microbiomes either as a cause or consequence
           ## of bad health. Might be worthwhile to consider plain
           ## weight or lenght as a proxy for age?! It'd be actually
           ## btter to use tail lenght for that, but not for now...
           bodyweight = Body_Weight,
           bodylength = Body_Length,
           BMI = residuals(lm(Body_Weight~Body_Length, data = .)),
           ## HI, the hybrid index. Hypothesis: "disturbed"
           ## microbiomes (especially mycobiomes) in hybrids as
           ## observed previously.
           HI = HI,
           ## Remark: All species interaction hypotheses
           ## (e.g. parasite infections, which would be available as
           ## derived sample variables) should be addressed as part of
           ## the response variable
           loc = Locality,
           batch = Batch) %>%
    full_join(aggregateDensity) %>%
  dplyr::select(mouseID, sex, lon, lat, year, bodyweight, bodylength, BMI, HI, loc, farm_ID, batch, Ntrapped, observed_arg, chao1_arg, shannon_arg, simpson_eve_arg,
                simpson_dom_arg)

colnames(allSData)

head(allSData)
```

### feed it back into the core data set after dropping the same NA mice as for the sample data

```{r}
PS.mOTUS.filt_jS <- subset_samples(PS.mOTUS.filt, !is.na(Body_Length))
PS.mOTUS.filt_jS@sam_data <- sample_data(allSData)

## check phyloseq object is correct
sample_names(PS.mOTUS.filt_jS@otu_table)
sample_names(PS.mOTUS.filt_jS@sam_data)

## make sure the sample names are the same for the sample_data and otu_table. We use the jS_mouseID column
sample_names(PS.mOTUS.filt_jS@sam_data) <- PS.mOTUS.filt_jS@sam_data$mouseID
sample_names(PS.mOTUS.filt_jS@otu_table) <- PS.mOTUS.filt_jS@sam_data$mouseID

```


Save processed OTU 
```{r}
saveRDS(PS.mOTUS.filt_jS, "./data/processed/PS.OTU.filt_jS.rds")
# PS.mOTUS_jS <- readRDS("./data/processed/PS.OTU_jS.rds")
```


### get only the selected variables
```{r}
## get selected variables
miceData <- unclass(sample_data(PS.mOTUS.filt_jS)) %>% 
  as_tibble() 

head(miceData)
nrow(miceData)
# [1] 858 mice
```

### check for missing coordinates in data
```{r}
summary(miceData)

miceData %>% 
  filter(is.na(lon))

head(miceData)
nrow(miceData)
```
No missing coordinates

### Save this data for later merging to the environmental covariates
```{r}

write.csv(miceData, "./data/raw/OTU_mice_data_filtered_gicquel_202411.csv")

```

## Exploring OTU

### load data
```{r}
# microbiome bacteria data
PS.OTU.filt_jS <- readRDS("./data/processed/PS.OTU.filt_jS.rds")
PS.OTU.filt_jS

#OTU
#View(otu_table(PS.OTU.filt_jS))

#mice
sample_names(PS.OTU.filt_jS)
sample_variables(PS.OTU.filt_jS)

#View(sample_data(PS.OTU.filt_jS))

#taxa
taxa_names(PS.OTU.filt_jS)
taxa_sums(PS.OTU.filt_jS)
tax_table(PS.OTU.filt_jS)

#View(tax_table(PS.OTU.filt_jS))

## filter data for bacteria
PS_bacteria <- subset_taxa(PS.OTU.filt_jS, Kingdom == "Bacteria")

#View(PS_bacteria@otu_table)

#View(PS_bacteria@tax_table)

## get data at family level
rank_names(PS_bacteria)[5]
PS_bacteria_fam <- tax_glom(PS_bacteria, taxrank = rank_names(PS_bacteria)[5])
PS_bacteria_fam

## 60 bacteria families in previous sample group
## 49 bacteria Families in 847 mice

## get data at Phylum level
rank_names(PS_bacteria)[2]
PS_bacteria_phyl <- tax_glom(PS_bacteria, taxrank = rank_names(PS_bacteria)[2])
PS_bacteria_phyl
## 12 bacteria phylum in 847 mice

## get data at Class level
rank_names(PS_bacteria)[3]
PS_bacteria_class <- tax_glom(PS_bacteria, taxrank = rank_names(PS_bacteria)[3])
PS_bacteria_class
## 23 bacteria classes in 847 mice

## get data at species level
rank_names(PS_bacteria)[7]
PS_bacteria_sp <- tax_glom(PS_bacteria, taxrank = rank_names(PS_bacteria)[7])
PS_bacteria_sp
## 146 bacteria species in 847 mice


```

49 bacteria families in the 858 mice (sample)
23 bacteria classes
12 bacteria phylums
146 bacteria species

# ARG data

## ARG.raw

Three mice individuals are filtered out:  
- AA_0266: 1 ARG identified
- SK_3209: 0 ARG identified
- AA_0444: 0 ARG identified

### Cleaning up NA in Taxa table
```{r}

PS.ARG.raw.filt 

# let's clean up genus column in the tax table

tax <- as.data.frame(tax_table(PS.ARG.raw.filt))

tax <- as.data.frame(tax_table(PS.ARG.raw.filt)) %>%
  mutate(
    AMR_gene_family = ifelse(str_detect(AMR_gene_family, "^NA$|^!RNA$"), NA, AMR_gene_family),
    Variant = ifelse(str_detect(Variant, "NA"), NA, Variant),
    Resistance_Mechanism = ifelse(str_detect(Resistance_Mechanism, "NA"), NA, Resistance_Mechanism),
    ARG = ifelse(str_detect(ARG, "^NA$|^!RNA$"), NA, ARG),
    ARG_CARD = ifelse(str_detect(ARG_CARD, "^NA$|^!RNA$"), NA, ARG_CARD),
    ARG_CARD_Short = ifelse(str_detect(ARG_CARD_Short, "^NA$|^!RNA$"), NA, ARG_CARD_Short)
  ) 


PS.ARG.raw.filt@tax_table <- tax_table(as.matrix(tax))

#View(PS.ARG.raw.filt@tax_table)

length(unique(tax_table(PS.ARG.raw.filt)[,"Variant"]))

```


### Prepare ARG

```{r prepare data}
## N mice trapped per location

aggregateDensity <- unclass(sample_data(PS.ARG.raw.filt)) %>%
    ## need to remove class "sample_data" (phyloseq)
    as_tibble() %>%
    ## need to remove NA the same data 
    filter(!is.na(Body_Length)
           #, !is.na(HI)
           ) %>% 
    ## creating Locality variable
    mutate(Locality = paste(round(Latitude,4), " ", round(Longitude,4))) %>%
    ## add total mice trapped per locality and year  
    group_by(Locality, Year) %>%
    ## Number of mice trapped at one locality (combined long/lat at
    ## 4-digit resolution, available in "Locality"). We processed all
    ## mice for each locality. We create the variable per year to
    ## estimate density at one point in time. Hypothesis: as measure
    ## of host population density this indicates increase transmission
    ## between mice. More host-associated transmissible taxa will be
    ## enriched. Potential alternatives not considered now:
    ## collapsing localities at larger scale (more rounding),
    ## persistent density over years for those)
    summarise(Ntrapped = n())  


## filter NAs and select variables
allSData <- unclass(sample_data(PS.ARG.raw.filt)) %>%
    ## need to remove class "sample_data" (phyloseq)
    as_tibble() %>%
    ## mice needing to be dropped because of NA values (might be
    ## worthwhile to imput this at some point, "Status" could be most
    filter(!is.na(Body_Length)
           #, !is.na(HI)
           ) %>% 
    ## add here other variable that could need filtering if included
    ## SELECTION FOR NOW:
    ## Latitude and longitude: the basic spatial coordinates, decimal
    ## format, rounded to 4 digits
    ## creating Locality variable
    mutate(Locality = paste(round(Latitude,4), " ", round(Longitude,4)),
           farm_ID = as.integer(factor(Locality))) %>%
    mutate(mouseID = Mouse_ID,
           sex = Sex, 
           lon = Longitude, 
           lat = Latitude, 
           ## year: the year of sampling, hypothesis: closer years are more
           ## similar as a result of community persistence or similarity of
           ## environmental condition in closer years.
           year = as.integer(Year),
           ## BMI: the residuals of body weight on body
           ## length. Hypothesis: mice with worse body condition have
           ## disturbed microbiomes either as a cause or consequence
           ## of bad health. Might be worthwhile to consider plain
           ## weight or lenght as a proxy for age?! It'd be actually
           ## btter to use tail lenght for that, but not for now...
           bodyweight = Body_Weight,
           bodylength = Body_Length,
           BMI = residuals(lm(Body_Weight~Body_Length, data = .)),
           ## HI, the hybrid index. Hypothesis: "disturbed"
           ## microbiomes (especially mycobiomes) in hybrids as
           ## observed previously.
           HI = HI,
           ## Remark: All species interaction hypotheses
           ## (e.g. parasite infections, which would be available as
           ## derived sample variables) should be addressed as part of
           ## the response variable
           loc = Locality,
           batch = Batch) %>%
    full_join(aggregateDensity) %>%
  dplyr::select(mouseID, sex, lon, lat, year, bodyweight, bodylength, BMI, HI, loc, farm_ID, batch, Ntrapped, observed_arg, chao1_arg, shannon_arg, simpson_eve_arg,
                simpson_dom_arg)

head(allSData)
```

### feed it back into the core data set after dropping the same NA mice as for the sample data
```{r}
PS.ARG.raw.filt_jS <- subset_samples(PS.ARG.raw.filt, !is.na(Body_Length))
PS.ARG.raw.filt_jS@sam_data <- sample_data(allSData)

## check phyloseq object is correct
sample_names(PS.ARG.raw.filt_jS@otu_table)
sample_names(PS.ARG.raw.filt_jS@sam_data)

## make sure the sample names are the same for the sample_data and otu_table. We use the jS_mouseID column
sample_names(PS.ARG.raw.filt_jS@sam_data)<- PS.ARG.raw.filt_jS@sam_data$mouseID
sample_names(PS.ARG.raw.filt_jS@otu_table)<- PS.ARG.raw.filt_jS@sam_data$mouseID

PS.ARG.raw.filt_jS

saveRDS(PS.ARG.raw.filt_jS, "./data/processed/PS.ARG.raw.filt_jS.rds")
```


### get only the selected variables
```{r}
## get selected variables
miceData <- unclass(sample_data(PS.ARG.raw.filt_jS)) %>% 
  as_tibble() 

head(miceData)
nrow(miceData)
# [1] 859 mice
```

### check for missing coordinates in data
```{r}
summary(miceData)

miceData %>% 
  filter(is.na(lon))

head(miceData)
nrow(miceData)
```
No missing coordinates

### Save this data for later merging to the environmental covariates
```{r}

write.csv(miceData, "./data/raw/ARG_mice_data_filtered_gicquel_202411.csv")

```


# ARG data - normalized

ARG normalized count

Three mice individuals are filtered out:  
- AA_0266: 1 ARG identified
- SK_3209: 0 ARG identified
- AA_0444: 0 ARG identified
- SK_3018: 1 ARG identified


### Cleaning up NA in Taxa table
```{r}

PS.ARG.FPKM.filt

# let's clean up genus column in the tax table

tax <- as.data.frame(tax_table(PS.ARG.FPKM.filt))

tax <- as.data.frame(tax_table(PS.ARG.FPKM.filt)) %>%
  mutate(
    AMR_gene_family = ifelse(str_detect(AMR_gene_family, "^NA$|^!RNA$"), NA, AMR_gene_family),
    Variant = ifelse(str_detect(Variant, "NA"), NA, Variant),
    Resistance_Mechanism = ifelse(str_detect(Resistance_Mechanism, "NA"), NA, Resistance_Mechanism),
    ARG = ifelse(str_detect(ARG, "^NA$|^!RNA$"), NA, ARG),
    ARG_CARD = ifelse(str_detect(ARG_CARD, "^NA$|^!RNA$"), NA, ARG_CARD),
    ARG_CARD_Short = ifelse(str_detect(ARG_CARD_Short, "^NA$|^!RNA$"), NA, ARG_CARD_Short)
  ) 


PS.ARG.FPKM.filt@tax_table <- tax_table(as.matrix(tax))

#View(PS.ARG.raw@tax_table)

length(unique(tax_table(PS.ARG.FPKM.filt)[,"Variant"]))
#81 variants

```


## Prepare ARG

```{r prepare data}
## N mice trapped per location

aggregateDensity <- unclass(sample_data(PS.ARG.FPKM.filt)) %>%
    ## need to remove class "sample_data" (phyloseq)
    as_tibble() %>%
    ## need to remove NA the same data 
    filter(!is.na(Body_Length)
           #, !is.na(HI)
           ) %>% 
    ## creating Locality variable
    mutate(Locality = paste(round(Latitude,4), " ", round(Longitude,4))) %>%
    ## add total mice trapped per locality and year  
    group_by(Locality, Year) %>%
    ## Number of mice trapped at one locality (combined long/lat at
    ## 4-digit resolution, available in "Locality"). We processed all
    ## mice for each locality. We create the variable per year to
    ## estimate density at one point in time. Hypothesis: as measure
    ## of host population density this indicates increase transmission
    ## between mice. More host-associated transmissible taxa will be
    ## enriched. Potential alternatives not considered now:
    ## collapsing localities at larger scale (more rounding),
    ## persistent density over years for those)
    summarise(Ntrapped = n())  


## filter NAs and select variables
allSData <- unclass(sample_data(PS.ARG.FPKM.filt)) %>%
    ## need to remove class "sample_data" (phyloseq)
    as_tibble() %>%
    ## mice needing to be dropped because of NA values (might be
    ## worthwhile to imput this at some point, "Status" could be most
    filter(!is.na(Body_Length)
           #, !is.na(HI)
           ) %>% 
    ## add here other variable that could need filtering if included
    ## SELECTION FOR NOW:
    ## Latitude and longitude: the basic spatial coordinates, decimal
    ## format, rounded to 4 digits
    ## creating Locality variable
    mutate(Locality = paste(round(Latitude,4), " ", round(Longitude,4))) %>%
    mutate(mouseID = Mouse_ID,
           sex = Sex, 
           lon = Longitude, 
           lat = Latitude, 
           ## year: the year of sampling, hypothesis: closer years are more
           ## similar as a result of community persistence or similarity of
           ## environmental condition in closer years.
           year = as.integer(Year),
           ## BMI: the residuals of body weight on body
           ## length. Hypothesis: mice with worse body condition have
           ## disturbed microbiomes either as a cause or consequence
           ## of bad health. Might be worthwhile to consider plain
           ## weight or lenght as a proxy for age?! It'd be actually
           ## btter to use tail lenght for that, but not for now...
           bodyweight = Body_Weight,
           bodylength = Body_Length,
           BMI = residuals(lm(Body_Weight~Body_Length, data = .)),
           ## HI, the hybrid index. Hypothesis: "disturbed"
           ## microbiomes (especially mycobiomes) in hybrids as
           ## observed previously.
           HI = HI,
           ## Remark: All species interaction hypotheses
           ## (e.g. parasite infections, which would be available as
           ## derived sample variables) should be addressed as part of
           ## the response variable
           loc = Locality,
           batch = Batch) %>%
    full_join(aggregateDensity) %>%
  dplyr::select(mouseID, sex, lon, lat, year, bodyweight, bodylength, BMI, HI, loc, batch, Ntrapped, observed_arg, chao1_arg, shannon_arg, simpson_eve_arg,
                simpson_dom_arg)

head(allSData)
```

### feed it back into the core data set after dropping the same NA mice as for the sample data
```{r}
PS.ARG.FPKM.filt_jS <- subset_samples(PS.ARG.FPKM.filt, !is.na(Body_Length))
PS.ARG.FPKM.filt_jS@sam_data <- sample_data(allSData)

## check phyloseq object is correct
sample_names(PS.ARG.FPKM.filt_jS@otu_table)
sample_names(PS.ARG.FPKM.filt_jS@sam_data)

## make sure the sample names are the same for the sample_data and otu_table. We use the jS_mouseID column
sample_names(PS.ARG.FPKM.filt_jS@sam_data)<- PS.ARG.FPKM.filt_jS@sam_data$mouseID
sample_names(PS.ARG.FPKM.filt_jS@otu_table)<- PS.ARG.FPKM.filt_jS@sam_data$mouseID

PS.ARG.FPKM.filt_jS

saveRDS(PS.ARG.FPKM.filt_jS, "./data/processed/PS.ARG.FPKM.filt_jS.rds")
```


### get only the selected variables
```{r}
## get selected variables
miceData <- unclass(sample_data(PS.ARG.FPKM.filt_jS)) %>% 
  as_tibble() 

head(miceData)
nrow(miceData)
# [1] 847 mice
```

### check for missing coordinates in data
```{r}
summary(miceData)

miceData %>% 
  filter(is.na(lon))

head(miceData)
nrow(miceData)
```
No missing coordinates


### Save this data for later merging to the environmental covariates
```{r}

write.csv(miceData, "./data/raw/ARG_norm_mice_data_filtered_gicquel_202411.csv")

```






