---
title: "4 - Mice ARGs - jSDM"
author: "Morgane Gicquel"
date: "2025-08-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*******************RUN PARALLEL IN CS1************************************
01.09.2024 This code is only for the IZW CS1 server, so the parallelizing works, until next server update.
```{r}
#install.packages("RhpcBLASctl")
library('RhpcBLASctl')
omp_set_num_threads(1)
blas_set_num_threads(1)
Sys.setenv(NUMEXPR_NUM_THREADS = 1,
           OPENBLAS_NUM_THREADS = 1,
           VECLIB_MAXIMUM_THREADS = 1,
           BLIS_NUM_THREADS = 1,
           OMP_NUM_THREADS = 1, 
           MKL_NUM_THREADS = 1)
```
***************************************************************************

# Model structure 

We have a total 340 different antibiotic resistant genes found in the samples.  

Response variable: presence / absence of each ARG in mice gut microbiome. 

Explanatory variables:  
- Mouse characteristics: sex, BMI   
- Mouse density: Ntrapped  
- Landcover: imperv, tcd, swf_perc, agri    
- Accessibility: dist_road, dist_path
- Farm animals: cat_dens, pig_dens, pou_dens  
- Climatic: precipitation_aug, air_temp_aug  

Random:  
- Spatial autocorrelation (longitude and latitude)
- Farm ID
- Year  

Traits:  
- Localization  
- Mobility  
- Drug class  
- Resistance mechanism


## Preparing workspace
```{r prepare workspace}
#setwd("/mnt/dep6/gicquel/mus_musculus_de_amr_gicquel_m/")

# Indicate date of the current run
date <- "2025-07-18"

# Loading packages needed, as well as custom ggplot themes
source("scripts/source_packages.R") # R libraries
source("scripts/source_themes_mg.R")

# Setting working and output directories 
wrkdir <- getwd()
model_wd <- paste0(wrkdir, "/output/jSDM_ARG_aro_pa_", date)
```

## Loading data

### Mice data with associated environmental covariates
```{r}
# mice related data
mice_data <- read.csv("./data/processed/Table_ARG_mice_envcov_100m_202507.csv")
head(mice_data)
table(mice_data$year)

# removing samples with missing information on sex of mice
mice_data <- mice_data %>%
  filter(sex != "") 

```

### Traits data

```{r}
# traits data
arg_trait <- read.csv("./data/raw/arg_mobility.csv")
head(arg_trait)

```


### ARGs data
Loading the data with ARGs community for each sample.

```{r}
# ARG bacteria data
PS.ARG.raw_jS <- readRDS("./data/processed/PS.ARG.raw.filt_jS.rds")
PS.ARG.raw_jS <- PS.ARG.raw_jS %>%
  subset_samples(., sex != "")

sample_data(PS.ARG.raw_jS@sam_data)

# Change order of taxa

new_tax <- data.frame(PS.ARG.raw_jS@tax_table) %>%
  mutate(ARO = rownames(.)) %>%
  dplyr::select(ARO, ARG_CARD_Short, AMR_gene_family, Resistance_Mechanism, Variant)
  
PS.ARG.raw_jS@tax_table <- tax_table(as.matrix(new_tax))

## get data at family level
PS_ARG_fam <- tax_glom(PS.ARG.raw_jS, taxrank = "ARG_CARD_Short")
PS_ARG_fam

#View(tax_table(PS_ARG_fam))

arg_card_short <- data.frame(PS_ARG_fam@tax_table) %>%
  dplyr::select(ARO, ARG_CARD_Short)
  

## 98 ARG AMR gene family in 846 mice

PS_ARG_mech <- tax_glom(PS.ARG.raw_jS, taxrank = "Resistance_Mechanism")
PS_ARG_mech

## 6 taxa in 858 mice

PS_ARG_var <- tax_glom(PS.ARG.raw_jS, taxrank = "Variant")
PS_ARG_var

unique(data.frame(PS_ARG_var@tax_table)$Variant)

## 48 taxa in 858 mice, but 37 unique ones

## get data at ARO level

PS_ARG_aro <- tax_glom(PS.ARG.raw_jS, taxrank = "ARO")
  
PS_ARG_aro

#View(tax_table(PS_ARG_aro))

unique(data.frame(PS_ARG_aro@tax_table)$ARO)

```

### Associating traits data with ARGs taxa matrix
```{r}
# traits summary

traits_summary <- arg_trait %>%
  left_join(., new_tax) %>%
  dplyr::select(-Variant) 

unique(traits_summary$drug_class)

traits_summary <- traits_summary %>% 
  mutate(drug_class_grp = case_when(
    drug_class %in% c("beta-lactam", "cephalosporin", "carbapenem") ~ "Beta-lactams",
    drug_class %in% c("macrolide", "lincosamide", "streptogramin", "elfamycin") ~ "MLS",
    drug_class %in% c("fluoroquinolone", "rifamycin", "phenicol") ~ "Aromatics",
    drug_class %in% c("pleuromutilin", "fusidane") ~ "Terpenoids",
    drug_class %in% c("aminoglycoside") ~ "Aminoglycosides",
    drug_class %in% c("sulfonamide") ~ "Sulfonamides",
    drug_class %in% c("glycopeptide", "peptide") ~ "Peptides",
    drug_class %in% c("tetracycline") ~ "Tetracyclines",
    drug_class %in% c("diaminopyrimidine","phosphonic_acid", "aminocoumarin",
                      "nitroimidazole", "nucleoside", "zoliflodacin","mupirocin") ~ "Other",
    drug_class == "multidrug" ~ "Multidrug",
    TRUE ~ "Unclassified"  # Catch any unexpected drug_class
  ),
  localization_grp = case_when(
    localization %in% c("multiple_mge","pathogenicity_islands") ~ "mge_pai",
    .default = localization)
  )


```

## Extract ARGs presence / absence data for each mouse
```{r get table with pa data}

## change the row names of tax table and col names of otu tables for new ones so they are easier to merge
rownames(PS_ARG_aro@tax_table) <- paste0("ASV", 1:nrow(PS_ARG_aro@tax_table))
colnames(PS_ARG_aro@otu_table) <- paste0("ASV", 1:ncol(PS_ARG_aro@otu_table))

PS_ARG_aro@otu_table
PS_ARG_aro@tax_table

otut <- data.frame(PS_ARG_aro@otu_table)
taxt <- data.frame(PS_ARG_aro@tax_table)
taxt$ARO

rownames(otut) <- make.names(taxt$ARO, unique = TRUE)

head(otut)

## make data presence / absence
otut_pa <- otut
otut_pa[otut_pa > 0] <- 1
 
head(otut_pa)

## need to set the mouseId as a column to merge the data later
otut_pa <- otut_pa %>%
  t(.) %>% 
  data.frame(.) %>%
  tibble::rownames_to_column("mouseID")

```


```{r check duplicates}
## to remove duplicates, we need to transpose the data, merge the rows with the same name and transpose again to the original format
sampleID <- rownames(otut_pa)
taxaID <- colnames(otut_pa)

length(unique(taxaID)) # 341: 340 Fam + mouseID
length(unique(sampleID)) # 846 samples
# No dup
```

## Prepare final mice data

Assign precipitation and temperature values to the year the mouse where captured.

```{r}
head(mice_data)

mice_data %>% 
  group_by(loc) %>% 
  # mutate(numId = seq(1:length(unique(loc)))) 
  mutate(locID =  cur_group_id()) %>% 
  dplyr::select(mouseID, locID) 

## filter environmnetal variables and set pig density NA to 0
mice_data2 <- mice_data %>% 
  group_by(loc) %>% 
  mutate(numlocID =  cur_group_id()) %>% 
  mutate(locID = paste0("loc_", numlocID)) %>% 
  ungroup() %>% 
  ungroup() %>%                       
  mutate(mouseID = paste0("ASV", 1:n())) %>%
  mutate(precipitation_aug = case_when(
    year == 2015 ~ prec_aug_2015,
    year == 2016 ~ prec_aug_2016,
    year == 2017 ~ prec_aug_2017,
    year == 2018 ~ prec_aug_2018,
    year == 2019 ~ prec_aug_2019,
    year == 2021 ~ prec_aug_2021,
    year == 2022 ~ prec_aug_2022,
    TRUE ~NA)) %>% 
    mutate(air_temp_aug = case_when(
    year == 2015 ~ temp_aug_2015,
    year == 2016 ~ temp_aug_2016,
    year == 2017 ~ temp_aug_2017,
    year == 2018 ~ temp_aug_2018,
    year == 2019 ~ temp_aug_2019,
    year == 2021 ~ temp_aug_2021,
    year == 2022 ~ temp_aug_2022,
    TRUE ~NA)) %>% 
  mutate(sex = as.factor(sex), 
         municipality_code = as.factor(as.character(gem_char)), 
         region = as.factor(region)) %>% 
  mutate(pig_dens = replace_na(pig_dens, 0)) %>% 
  dplyr::select(-c(loc, prec_aug_2015, prec_aug_2016, prec_aug_2017, prec_aug_2018, prec_aug_2019, prec_aug_2021, prec_aug_2022, 
                   temp_aug_2015, temp_aug_2016, temp_aug_2017, temp_aug_2018, temp_aug_2019, temp_aug_2021, temp_aug_2022,
                   ID, CATO, PITO, POTO, gem_char, HI))
mice_data2

## add bacteria
all_data <- mice_data2 %>% 
  left_join(otut_pa, by = c("mouseID"))

head(all_data)
```

# Run jSDM

For the JSDM we need:
- a matrix of species: sites in rows (mice) and species in columns (OTU)
- a data frame with the environmental covariates for sites: sites (mice) in rows, same order as above
- a trait data frame with information on the species (OTU/ARO)
- random structure
- spatial structure

## Set up variables and model structure

```{r}
colnames(all_data)

# response matrix 
my_response <- colnames(otut_pa)[-1] ## remove mouseID
response_pa <- all_data %>% 
  dplyr::select(all_of(my_response)) %>% 
  as.matrix()

# explanatory df
my_cov <- c("sex", "BMI",  
            "Ntrapped", 
            "agri", "imperv", "tcd", "swf", 
            "dist_road", "dist_path",
            "cat_dens", "pig_dens", "pou_dens", 
            "precipitation_aug", "air_temp_aug")

explanatory_var <- all_data %>% 
  dplyr::select(all_of(my_cov)) 

as.data.frame(explanatory_var[!complete.cases(explanatory_var),])

## spatial data (without duplicates)
xyData <- all_data %>%
  group_by(locID) %>% 
  summarise(lon = unique(lon), 
            lat = unique(lat)) %>% 
  tibble::column_to_rownames("locID")

## three nested random effects
studyDesign <- data.frame(sample = as.factor(all_data$mouseID), 
                          farm_ID = as.factor(all_data$locID), 
                          spatial_autocorrelation = as.factor(all_data$locID), 
                          year = as.factor(all_data$year))

rL3 <- HmscRandomLevel(units = levels(all_data$year)) # account for temporal correlation
rL2 <- HmscRandomLevel(units = levels(all_data$locID)) # accounts for nested sampling 
rL1 <- HmscRandomLevel(sData = xyData, units = levels(all_data$locID)) # accounts for spatial locations


## trait data

TrData_aro <- traits_summary %>%
  mutate(ARO = str_replace_all(ARO, c(":"),c("."))) %>%
  `rownames<-`(.[,1]) %>%
  dplyr::select(-ARO, -ARG_CARD_Short, -AMR_gene_family, -drug_class)

head(TrData_aro)

TrData_aro <- TrData_aro[match(colnames(response_pa), rownames(TrData_aro)), ]

unique(TrData_aro$drug_class_grp)

unique(TrData_aro$Resistance_Mechanism)

```


### Define MCMC parameters
```{r}
## Initial parameters to test the data
thin <- 10
samples <- 1000
transient <- 100
nChains <- 4
verbose <- 10

# Regression formula with all environmental covariates

XFormula_OTU = ~ sex + BMI + Ntrapped + 
  agri + imperv + tcd + swf + 
  dist_road + dist_path + 
  cat_dens + pig_dens + pou_dens + 
  precipitation_aug + air_temp_aug

## no traits or phylogeny are used at this time

TrFormula_aro = ~ localization_grp + mobility + drug_class_grp + Resistance_Mechanism


```


### Fit models for PRESENCE/ABSENCE  data 

*BINOMIAL DISTRIBUTION* ~> PROBIT MODEL

```{r}

m <- Hmsc(Y = response_pa, 
          XData = explanatory_var, 
          XFormula = XFormula_OTU,
          TrData = TrData_aro,
          TrFormula = TrFormula_aro,
          studyDesign = studyDesign, 
          ranLevels=list(
            "spatial_autocorrelation"= rL1, 
            "farm_ID" = rL2, 
            "year" = rL3),
          distr = "probit")

```

## Run model
```{r}
# Run Markov Chains
# test
m <- sampleMcmc(m, thin = 1, samples = 10)

tic("Hmsc model")
# the real model
m <- sampleMcmc(m, thin = thin, samples = samples, transient = transient, 
                nChains = nChains, verbose = verbose, nParallel = nChains)
toc()


# Set a name for the model
filename <- paste0(model_wd, "/jsdm_model_",date,".rds")

#Save the model
saveRDS(m, file = filename) 
# 
```

# Check the model

```{r model convergence}
#model path
filename <- paste0(model_wd, "/jsdm_model_",date,".rds")

# retrieve model
m <- readRDS(filename)

m
```

Hmsc object with 846 sampling units, 340 species, 15 covariates, 22 traits and 3 random levels
Posterior MCMC sampling with 4 chains each with 1000 samples, thin 10 and transient 100  

## Model convergence 

We evaluate MCMC convergence in terms of two kinds of parameters that we are especially interested in:
the species niches Beta, and the residual species associations Omega. 

Evaluate convergence: Effective sample size and Gelman-Rubin diagnostic (potential reduction factor).   
One can then reason that we should run our chains until the value of R is close to 1 (< 1.1).

```{r model convergence}
mpost <- convertToCodaObject(m)

saveRDS(mpost, paste0(model_wd, "/jsdm_mpost_coda_", date, ".rds"))

# Save plots of the chains
MCMCtrace(mpost$Beta, 
          pdf = TRUE, 
          open_pdf = FALSE,
          filename = paste0("jsdm_MCMC_trace_beta_",date),
          wd = model_wd)

MCMCtrace(mpost$Omega[[1]], 
          pdf = TRUE, 
          open_pdf = FALSE,
          filename = paste0("jsdm_MCMC_trace_omega_",date),
          wd = model_wd)

MCMCtrace(mpost$Gamma, 
          pdf = TRUE, 
          open_pdf = FALSE,
          filename = paste0("jsdm_MCMC_trace_gamma_",date),
          wd = model_wd)

par(mfrow=c(1,1))

# Numerical output
ess.beta <- effectiveSize(mpost$Beta)
ess.omega <- effectiveSize(mpost$Omega[[1]])
ess.gamma <- effectiveSize(mpost$Gamma)

gd.beta <- gelman.diag(mpost$Beta, multivariate = FALSE)$psrf
#gd.omega <- gelman.diag(mpost$Omega[[1]], multivariate = FALSE)$psrf
gd.gamma <- gelman.diag(mpost$Gamma, multivariate = FALSE)$psrf

# mean values
mean(ess.beta) #[1] 1623.94
mean(ess.omega) # [1] 835.2532
mean(ess.gamma) # [1] 927.4198

mean(gd.beta) #[1] 1.028207
#mean(gd.omega) #[1] impossible to compute
mean(gd.gamma) #[1] 1.036653

ess.beta_hist <- data.frame(ess.beta) %>%
  ggplot(aes(x=ess.beta)) + 
  geom_histogram(color="black", fill="white") +
  geom_vline(xintercept = mean(ess.beta), colour = "#440154FF", linewidth = 1, linetype = "dashed") +
  geom_text(aes(y = 400, x = 2500), label = round(mean(ess.beta),2), colour = "#440154FF", size = 3, family = "Poppins") +
  labs(x = "Beta ESS", y = "Count") +
  theme_mg()

ess.omega_hist <- data.frame(ess.omega) %>%
  ggplot(aes(x=ess.omega)) + 
  geom_histogram(color="black", fill="white") +
  geom_vline(xintercept = mean(ess.omega), colour = "#440154FF", linewidth = 1, linetype = "dashed") +
  geom_text(aes(y = 13500, x = 1500), label = round(mean(ess.omega),2), colour = "#440154FF", size = 3, family = "Poppins") +
  labs(x = "Omega ESS", y = "Count") +
  theme_mg()

ess.gamma_hist <- data.frame(ess.gamma) %>%
  ggplot(aes(x=ess.gamma)) + 
  geom_histogram(color="black", fill="white") +
  geom_vline(xintercept = mean(ess.gamma), colour = "#440154FF", linewidth = 1, linetype = "dashed") +
  geom_text(aes(y = 40, x = 1500), label = round(mean(ess.gamma),2), colour = "#440154FF", size = 3, family = "Poppins") +
  labs(x = "Gamma ESS", y = "Count") +
  theme_mg()

#
gd.beta_hist <- data.frame(gd.beta) %>%
  ggplot(aes(x=Point.est.)) + 
  geom_histogram(color="black", fill="white") +
  geom_vline(xintercept = mean(data.frame(gd.beta)$Point.est.), colour = "#440154FF", linewidth = 1, linetype = "dashed") +
  geom_text(aes(y = 3000, x = 1.1), label = round(mean(gd.beta),2), colour = "#440154FF", size = 3, family = "Poppins") +
  labs(x = "Beta GR", y = "Count") +
  theme_mg()

# gd.omega_hist <- data.frame(gd.omega) %>%
#   ggplot(aes(x=Point.est.)) + 
#   geom_histogram(color="black", fill="white") +
#   geom_vline(xintercept = mean(data.frame(gd.omega)$Point.est.), colour = "#440154FF", linewidth = 1, linetype = "dashed") +
#   geom_text(aes(y = 10, x = 1.01), label = round(mean(data.frame(gd.omega)$Point.est.),2), colour = "#440154FF", size = 3, family = "Poppins") +
#   labs(x = "Gelman-Rubin Diagnostic Omega", y = "Count") +
#   theme_mg()

gd.gamma_hist <- data.frame(gd.gamma) %>%
  ggplot(aes(x=Point.est.)) + 
  geom_histogram(color="black", fill="white") +
  geom_vline(xintercept = mean(data.frame(gd.gamma)$Point.est.), colour = "#440154FF", linewidth = 1, linetype = "dashed") +
  geom_text(aes(y = 150, x = 1.08), label = round(mean(gd.gamma),2), colour = "#440154FF", size = 3, family = "Poppins") +
  labs(x = "Gamma GR", y = "Count") +
  theme_mg()


PLOT_Model_convergence <- plot_grid(ess.beta_hist, 
                                    ess.omega_hist, 
                                    ess.gamma_hist,
                                    gd.beta_hist, 
                                    #gd.omega_hist,
                                    NA,
                                    gd.gamma_hist)
PLOT_Model_convergence

ggsave(paste0(model_wd, "/Plot_jsdm_convergence_hist_", date, ".png"), PLOT_Model_convergence, 
      dpi = 600, width = 32, height = 13.33, units = "cm")

#Save plot appendices

ggsave(filename = here("figures", "appendices", paste0("Plot_jsdm_convergence_hist_", date, ".png")), PLOT_Model_convergence, dpi = 600, width = 16, height = 12, units = "cm")


```

Model convergence:  
Average effective sample sizes are high (> 800), showing that the model estimates robust parameters, and indicate low autocorrelation and good mixing.  
Gelman-Rubin Diagnostic values are acceptable for betas (1.03) and gammas (1.04), which means chains have converged well. It couldn't be compiled for omegas for structural a reasons (huge matrices for each sample).  

```{r}

mpost <- readRDS(paste0(model_wd, "/jsdm_mpost_coda_", date, ".rds"))

```

## Model Fit (R2)

```{r model fit}
# Explanatory R2. Get predictions for the observed values 
preds <- computePredictedValues(m, expected = TRUE)

preds.values <- apply(abind(preds,along=3),c(1,2), median) # Median of the predictions
write.csv(preds.values, paste0(model_wd, "/jsdm_predicted_values_median_", date, ".csv"))
preds.values.mean <- apply(abind(preds, along = 3), c (1,2), mean) # Mean of the predictions
write.csv(preds.values.mean, paste0(model_wd, "/jsdm_predicted_values_mean_", date, ".csv"))

# R2 with the built in function
modelr2.explanatory <- evaluateModelFit(hM = m, predY = preds)
modelr2.explanatory
```

```{r}
# R2 of the model
mean(modelr2.explanatory$TjurR2, na.rm=T)
# [1] 0.1238726

# AUC of the model
mean(modelr2.explanatory$AUC, na.rm=T)
# [1] 0.9017532

# RMSE of the model
mean(modelr2.explanatory$RMSE, na.rm=T)
# [1] 0.1688351


temp.r2 <- cbind.data.frame(r2 = as.numeric(modelr2.explanatory$TjurR2), 
                            species = as.character(m$spNames))

ggplot(data = temp.r2, aes(y=r2, x=species))+
  geom_col() + 
  theme_mg() +
  theme(axis.text.x = element_text(angle = 45, hjust= 1, vjust = 1))

write.csv(modelr2.explanatory, paste0(model_wd, "/jsdm_explanatory_R2_", date, ".csv"), row.names = FALSE)

# R2 Manually comparing observed vs predicted
R2.sp <- matrix(NA, m$ns, 1)
for (i in 1:m$ns) {
  R2.sp[i, ] <- cor(preds.values.mean[, i],m$Y[, i])^2
}

mean(R2.sp, na.rm=TRUE)
# [1] 0.1746014

# Plot of the default R2
ggplot(data = as.data.frame(modelr2.explanatory$TjurR2), aes(y=modelr2.explanatory$TjurR2, x=temp.r2$species))+
  geom_col() + 
  xlab("Species") +
  ylab("R2 for each species") +
  ggtitle(paste0("ARG Resistance mechanism in mice - Default R2 species \n R2 = ", 
                   round(mean(modelr2.explanatory$R2, na.rm = TRUE), 2))) +
  theme_mg() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

ggsave(paste0(model_wd, "/Plot_jsdm_R2_species_default", date, ".png"))


# Plot of the manual R2
ggplot(data = as.data.frame(R2.sp), aes(y=V1, x=temp.r2$species))+
  geom_col() +
  xlab("Species") +
  ylab("R2 for each species") +
  ggtitle(paste0("ARG Resistance mechanism in mice - Obs vs pred R2 species \n R2 = ", 
                   round(mean(R2.sp, na.rm = TRUE), 2))) +
  theme_mg() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

ggsave(paste0(model_wd, "/Plot_jsdm_R2_species_manual", date, ".png"))
```

```{r site R2}
R2.site <- matrix(NA, m$ny, 1)
for (i in 1:m$ny) {
  R2.site[i, ] <- cor(preds.values.mean[i, ], m$Y[i, ])^2
}
mean(R2.site, na.rm = TRUE)
# [1] 0.7230648
```

Model fit:  
PseudoR2 (Tjur R2) = 0.13  
-> A value of 0.13 is relatively low, suggesting that the fixed effects only explain 13% of the variation in species occurrence probabilities.  
AUC = 0.91   
-> A value of 0.91 indicates strong predictive performance of the model. This suggests the model is quite effective in identifying sites with high or low probabilities of species occurrence.  
R2.sp = 0.17  
->  A value of 0.17 is moderate, suggesting that the model explains 17% of the variation in species-specific responses. This may indicates that many species are rare, or poorly predicted by environmental predictors included in the model.  
R2.site = 0.72   
-> A value of 0.72 is high, indicating that the model does a very good job at capturing site-level differences, showing that random effects, such as spatial structure, location, and so on are very important.  


# Model results

## Inspect Beta coefficients

The Beta coefficients represent the species-environment relationships, and we can visualize these relationships for different species and environmental variables.

```{r betas}
Beta.results <- as.data.frame(MCMCsummary(mpost$Beta))
write.csv(Beta.results, paste0(model_wd, "/jsdm_beta_coefficients_", date, ".csv"), row.names = TRUE)

postBeta <- getPostEstimate(m, parName = "Beta")
saveRDS(postBeta, file=paste0(model_wd, "/jsdm_postBeta_", date, ".rds"))

#postBeta <- readRDS(paste0(model_wd, "/jsdm_postBeta_", date, ".rds"))

#Get species and covariate names from model
spNames <- m$spNames
covNames <- m$covNames

#pull the data from postBeta for plotting
mbeta <- postBeta$mean
betaP <- postBeta$support

# Support 95%
#Choose the support level for Beta plot
supportLevel <- 0.95

sign <- sign(mbeta)

Beta_association_values <- data.frame(((betaP > supportLevel) + (betaP < (1 - supportLevel)) > 0) * sign) 
colnames(Beta_association_values) <- spNames
rownames(Beta_association_values) <- covNames
#Beta_association_values[Beta_association_values == 0] <- NA
  
PLOT_Betas_ARG_resmech_95 <- ggcorrplot(Beta_association_values, method = "square") +
  labs(fill = "") +
  theme_mg_cor() 

PLOT_Betas_ARG_resmech_95

ggsave(paste0(model_wd, "/Plot_jsdm_beta_sign95_", date, ".png"), PLOT_Betas_ARG_resmech_95, dpi = 600)

# Support 75%
#Choose the support level for Beta plot
supportLevel <- 0.75

sign <- sign(mbeta)

Beta_association_values <- data.frame(((betaP > supportLevel) + (betaP < (1 - supportLevel)) > 0) * sign) 
colnames(Beta_association_values) <- spNames
rownames(Beta_association_values) <- covNames
#Beta_association_values[Beta_association_values == 0] <- NA
  
PLOT_Betas_ARG_resmech_75 <- ggcorrplot(Beta_association_values, method = "square") +
  labs(fill = "") +
  theme_mg_cor() 

PLOT_Betas_ARG_resmech_75

ggsave(paste0(model_wd, "/Plot_jsdm_beta_sign75_", date, ".png"), PLOT_Betas_ARG_resmech_75, dpi = 600)

```


```{r betas_faceted_plot}
# Extracting Mech names and Beta summaries
ARG <- colnames(m$Y)

# Summary statistics from Beta
beta_summary <- summary(mpost$Beta)

# Creating model frame for plotting
modelFrame_plot <- data.frame(
  Variable = colnames(mpost$Beta[[1]]),
  Coefficient = beta_summary$statistics[, 1],
  CI_low = beta_summary$quantiles[, 1],
  Q_25 = beta_summary$quantiles[, 2],
  Q_50 = beta_summary$quantiles[, 3],
  Q_75 = beta_summary$quantiles[, 4],
  CI_high = beta_summary$quantiles[, 5]) %>%
  mutate(ARG = gsub( "\\.", " ", rep(ARG, each = 15)), # Add AMR and cleaned variables
         Variable = stringr::str_sub(stringr::word(Variable, 1), 3, -1), 
         significant = ifelse(CI_low > 0 | CI_high < 0, "Yes", "No")) # Flag significant estimates (credible intervals not overlapping zero)

write.csv(modelFrame_plot, paste0(model_wd, "/jsdm_beta_modelFrame_", date, ".csv"), row.names = FALSE)


```

```{r betas_ResMech_individual_plots}

# Loop through each unique Resistance Mechanism
for (ARG_lab in unique(modelFrame_plot$ARG)) {
  
  # Filter the data for the current Resistance Mechanism
  filtered_data <- modelFrame_plot %>% filter(ARG == ARG_lab)
  
  # Create the plot for the current ResMech
  myplot <- ggplot(filtered_data, aes(x = Variable, y = Coefficient, colour = Variable)) + 
    geom_hline(yintercept = 0, colour = gray(1/2), linetype = "dashed") + 
    geom_linerange(aes(ymin = CI_low, ymax = CI_high, fill = significant),
                   size = 0.8, position = position_dodge(width = 1.2/2)) + 
    geom_linerange(aes(ymin = Q_25, ymax = Q_75, fill = significant),
                   size = 1.5, position = position_dodge(width = 1.2/2)) + 
    geom_pointrange(aes(y = Coefficient, ymin = Q_25, ymax = Q_75, fill = significant),
                    size = 0.5, shape = 21, position = position_dodge(width = 1.2/2)) +
    scale_fill_manual(values = c("White", "Black")) +
    scale_x_discrete(limits = rev(unique(modelFrame_plot$Variable))) +
    coord_flip() + 
    scale_colour_viridis_d(option = "viridis", begin = 0.1, end = 0.9, guide = guide_legend(reverse = TRUE)) +
    theme_mg() +
    theme(
       panel.background = element_rect(fill = NA),
       panel.grid.major = element_blank(), 
       axis.line = element_line(colour = "black"),
       legend.position = "none") +
    labs(title = paste0("Effects of variables on ", ARG),
         x = "Variables",
         y = "Coefficient Estimate")
  
  # Define the file name for the plot
  plotfile <- paste0(model_wd, "/BetaPlots/ARG/Plot_betas_", gsub(" ", "_", ARG_lab), "_", date, ".png")
  
  # Save the plot as a PNG file
  ggsave(plot = myplot, filename = plotfile, height = 13.33, width = 15.79, units="cm")
}

```

```{r betas_effects_individual_plots}

# Plotting and saving the effects of the variables
for (i in 1:length(unique(modelFrame_plot$Variable))){
  var <- unique(modelFrame_plot$Variable)[i]
  
  myplot <- modelFrame_plot %>% 
    filter(Variable == var) %>% 
  ggplot(aes(group = ARG, colour = significant)) + 
  geom_hline(yintercept = 0, colour = gray(1/2), lty = 2) + 
  geom_linerange(aes(x = ARG, ymin = CI_low,
                                ymax = CI_high, fill = significant),
                            lwd = 0.8, position = position_dodge(width = 1.2/2)) + 
  geom_linerange(aes(x = ARG, ymin = Q_25,
                                ymax = Q_75, fill = significant),
                            lwd = 1.5, position = position_dodge(width = 1.2/2)) + 
  geom_pointrange(aes(x = ARG, y = Coefficient, ymin = Q_25,
                                 ymax = Q_75, fill = significant),
                             lwd = 1/2, shape = 21, position = position_dodge(width = 1.2/2)) +
  scale_fill_manual(values = c("White", "black"))+
  
  # scale_y_continuous(limits = c(-4.2, 3)) +
  coord_flip() + 
  scale_colour_viridis_d(option = "viridis", begin = 0.1, end = 0.9, 
                         guide = guide_legend(reverse = TRUE)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_blank(), 
    axis.line = element_line(colour = "black"),
    legend.position = "NONE") +
    ggtitle(paste0("Effects of ", var))
  
  plotfile <- paste0(model_wd, "/BetaPlots/EnvCov/Plot_betas_", var, "_", date, ".png")
  ggsave(plot = myplot, plotfile, height = 16, width = 20)
  
}

```


## Inspect Gamma coefficients

The Gamma coefficients represent the species traits-environment relationships, and we can visualize these relationships for different species and environmental variables.

```{r gammas}
Gamma.results <- as.data.frame(MCMCsummary(mpost$Gamma))
write.csv(Gamma.results, paste0(model_wd, "/jsdm_Gamma_coefficients_", date, ".csv"), row.names = TRUE)

postGamma <- getPostEstimate(m, parName = "Gamma")
saveRDS(postGamma, file=paste0(model_wd, "/jsdm_postGamma_", date, ".rds"))

#postGamma <- readRDS(paste0(model_wd, "/jsdm_postGamma_", date, ".rds"))

#Get species and covariate names from model
trNames <- m$trNames
covNames <- m$covNames

#pull the data from postGamma for plotting
mGamma <- postGamma$mean
GammaP <- postGamma$support


Gamma_association_mean <- data.frame(mGamma)
colnames(Gamma_association_mean) <- trNames
rownames(Gamma_association_mean) <- covNames
#Gamma_association_values[Gamma_association_values == 0] <- NA

Gamma_association_support <- data.frame(GammaP)
colnames(Gamma_association_support) <- trNames
rownames(Gamma_association_support) <- covNames

Gamma_association_values_df <- Gamma_association_mean %>%
  mutate(covNames = rownames(Gamma_association_mean)) %>%
  gather(trait,value,-covNames) %>%
  left_join(., Gamma_association_support %>%
  mutate(covNames = rownames(Gamma_association_support)) %>%
  gather(trait, support_lvl, -covNames)  %>%
  mutate(sign = ifelse(support_lvl > 0.95 | support_lvl < (1 - 0.95), "*", 
                       ifelse(support_lvl > 0.75 | support_lvl < (1 - 0.75), "·", NA)))) %>%
  mutate(type = str_replace_all(trait, c(".*localization.*" = "Localization",
                                         ".*mobility.*" = "Mobility",
                                         ".*drug_class.*" = "Drug class",
                                         ".*Resistance_Mechanism.*" = "Resistance mechanism",
                                         ".*(Intercept).*" = "Intercept")),
         trait = str_replace_all(trait, c("localization_grp" = "",
                                         "mobility" = "",
                                         "drug_class_grp" = "",
                                         "Resistance_Mechanism" = "",
                                         "antibiotic " = "",
                                         " to antibiotic" = "",
                                         ".*Intercept.*" = "Intercept",
                                         "TRUE" = "mobile"))) %>%
  mutate(covNames = factor(covNames, levels = c("(Intercept)", "sexM", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precipitation_aug", "air_temp_aug"),
                           labels = c("(Intercept)", "sex-M", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precip", "temp")),
         type = factor(type, levels = c("Intercept","Mobility","Localization","Drug class","Resistance mechanism"), labels = c("Intercept","Mobility","Localization","Drug class","Resistance\nmechanism"))) 

#unique(Gamma_association_values_df$covNames)
#unique(Gamma_association_values_df$type)
library(ggforce)

### sample size

# N ARGs 
# N mice
# N farm ID

data_ARG <- psmelt(PS_ARG_aro) %>%
  left_join(traits_summary) 

mobile <- data_ARG %>%
  group_by(mobility) %>%
  filter(Abundance > 0) %>%
  summarize(N_ARG = length(unique(ARO)),
            N_mice = length(unique(mouseID)),
            N_farmID = length(unique(farm_ID))) %>%
  rename(trait = mobility)
  
localization <- data_ARG %>%
  group_by(localization_grp) %>%
  filter(Abundance > 0) %>%
  summarize(N_ARG = length(unique(ARO)),
            N_mice = length(unique(mouseID)),
            N_farmID = length(unique(farm_ID))) %>%
  rename(trait = localization_grp)

drug_class <- data_ARG %>%
  group_by(drug_class_grp) %>%
  filter(Abundance > 0) %>%
  summarize(N_ARG = length(unique(ARO)),
            N_mice = length(unique(mouseID)),
            N_farmID = length(unique(farm_ID))) %>%
  rename(trait = drug_class_grp)

res_mech <- data_ARG %>%
  group_by(Resistance_Mechanism) %>%
  filter(Abundance > 0) %>%
  summarize(N_ARG = length(unique(ARO)),
            N_mice = length(unique(mouseID)),
            N_farmID = length(unique(farm_ID))) %>%
  rename(trait = Resistance_Mechanism)

intercept <-  data_ARG %>%
  filter(mobility == "FALSE",
         localization_grp == "chromosome", 
         drug_class_grp == "Aminoglycosides", 
         Resistance_Mechanism == "antibiotic efflux") %>% 
  filter(Abundance > 0) %>%
  summarize(N_ARG = length(unique(ARO)),
            N_mice = length(unique(mouseID)),
            N_farmID = length(unique(farm_ID)))


N_arg <- mobile %>%
  rbind(., localization) %>%
  rbind(., drug_class) %>%
  rbind(., res_mech) %>%
  mutate(trait = str_replace_all(trait, c("antibiotic " = "",
                                          " to antibiotic" = "",
                                          "TRUE" = "mobile"))) %>%
  filter(!trait %in% c("FALSE", "chromosome", "Aminoglycosides", "efflux")) %>%
  left_join(Gamma_association_values_df) %>%
  filter(covNames != "(Intercept)") %>%
  dplyr::select(-covNames, -value, -support_lvl, -sign) %>%
  gather("covNames","sample", 2:4) %>%
  distinct() %>%
  mutate(covNames = str_replace_all(covNames, c("N_ARG" = "# ARG",
                                          "N_mice" = "# mice",
                                          "N_farmID" = "# farm")),
         value = NA,
         support_lvl = NA,
         sign = NA)

# Plotting

PLOT_Gamma_all <- Gamma_association_values_df %>%
  full_join(N_arg) %>%
  mutate(covNames = factor(covNames, levels = c("sex-M", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precip", "temp", "# ARG", "# mice","# farm"))) %>%
  mutate(group = case_when(covNames %in% c("sex-M", "BMI") ~ "Mouse\ncharacteristics",
                           covNames == c("Ntrapped") ~ "Mouse\ndensity",
                           covNames %in% c("agri", "imperv", "tcd", "swf") ~ "Landcover",
                           covNames %in% c("dist_road", "dist_path") ~ "Accessibility",
                           covNames %in% c("cat_dens", "pig_dens", "pou_dens") ~ "Farm animal\ndensity",
                           covNames %in% c("precip", "temp") ~ "Climatic",
                           covNames %in% c("# ARG", "# mice","# farm") ~ "Sample\nsize"),
         group = factor(group, levels = c("Mouse\ncharacteristics","Mouse\ndensity","Landcover","Accessibility","Farm animal\ndensity","Climatic","Sample\nsize"))) %>%
  filter(covNames != "(Intercept)") %>%
  ggplot(., aes(y = trait, x = covNames, fill = value)) +
  geom_tile(colour = "black") +
  geom_text(aes(label = sign), size = 5, colour = "black") +
  geom_text(aes(label = sample), size = 2.5, family = "Poppins") +
  #scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-4.5, 10.5)) +
  scale_fill_gradientn(colors = c("blue4", "blue3","white", "red3","red4"),
                       values = scales::rescale(c(-5,-2, 0, 5, 10)),
                       breaks = c(-5,0,5,10),
                       limits = c(-5,10.5),
                       na.value="grey99") +
  scale_x_discrete(position = "top", guide = guide_axis(angle = 90)) +
#  facet_col(type ~ group, scales = "free_x",  space = "free", strip.position = c("top", "right")) +
  facet_grid(type ~ group, scales = "free", space = "free", switch = "y") +
  labs(fill = "Coefficient estimate\n(gamma)") +
  theme_mg_facets() +
  theme(
    text = element_text(size = 8),
    axis.text.y = element_text(hjust = 1, vjust = 0.4, size = 8),
    axis.text.x = element_text(size = 8),
    axis.title = element_blank(),
    axis.line = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    strip.placement = "outside",
    strip.background = element_rect(fill = "grey90", colour = "grey"),
    strip.text.x = element_text(angle = 90, size = 8),
    strip.text.y.left = element_text(angle = 0, size = 8),
    panel.spacing = unit(0, "cm"))

PLOT_Gamma_all

# Save the plot as a PNG

ggsave(PLOT_Gamma_all , filename = here("figures", paste0("Plot_jsdm_pa_traits_gammas_", date, ".png")), height = 20, width = 16, units = "cm", dpi = 600)

# Save the plot as a SVG

ggsave(PLOT_Gamma_all , filename = here("figures", paste0("Plot_jsdm_pa_traits_gammas_", date, ".svg")), height = 20, width = 16, units = "cm", dpi = 600)

```


```{r model_frame, echo=FALSE, fig.height=3, fig.width=6.3}

mpost <- readRDS(paste0(model_wd, "/jsdm_mpost_coda_", date, ".rds"))

gamma_summary <- summary(mpost$Gamma)

Traits <- trNames

# Creating model frame for plotting
modelFrame_plot <- data.frame(
  Variable = colnames(mpost$Gamma[[1]]),
  Coefficient = gamma_summary$statistics[, 1],
  CI_low = gamma_summary$quantiles[, 1],
  Q_25 = gamma_summary$quantiles[, 2],
  Q_50 = gamma_summary$quantiles[, 3],
  Q_75 = gamma_summary$quantiles[, 4],
  CI_high = gamma_summary$quantiles[, 5]) %>%
  mutate(trait = rep(Traits, each=15), # Add traits and cleaned variables
         Variable = stringr::str_sub(stringr::word(Variable, 1), 3, -1)) %>% # Flag significant estimates (credible intervals not overlapping zero)
  mutate(type = str_replace_all(trait, c(".*localization.*" = "Localization",
                                         ".*mobility.*" = "Mobility",
                                         ".*drug_class.*" = "Drug class",
                                         ".*Resistance_Mechanism.*" = "Resistance mechanism",
                                         ".*(Intercept).*" = "Reference")),
         trait = str_replace_all(trait, c("localization_grp" = "",
                                         "mobility" = "",
                                         "drug_class_grp" = "",
                                         "Resistance_Mechanism" = "",
                                         "antibiotic " = "",
                                         " to antibiotic" = "",
                                         ".*Intercept.*" = "Reference",
                                         "TRUE" = "mobile"))) %>%
  mutate(Variable = factor(Variable, levels = c("(Intercept)", "sexM", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precipitation_aug", "air_temp_aug"),
                           labels = c("(Intercept)", "sex-M", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precip", "temp")),
         type = factor(type, levels = c("Reference","Mobility","Localization","Drug class","Resistance mechanism")))  %>%
  left_join(., Gamma_association_values_df, by = c("Variable" = "covNames", "type", "trait")) %>%
  mutate(significant = ifelse(sign %in% c("*","·"), "Yes", "No"))

filtered_df <- modelFrame_plot %>%
  filter(type != "Reference", Variable != "(Intercept)")  %>%
  droplevels()

```

```{r traits_individual_gamma_plots, echo=FALSE, fig.height=3, fig.width=6.3}

for (i in 1:length(unique(filtered_df$type))){
  type_lab <- unique(filtered_df$type)[i]

Plot_gamma <- filtered_df %>% 
  filter(type == type_lab) %>%
  ggplot(aes(x = Variable, colour = trait)) + 
  scale_x_discrete(limits = rev(levels(filtered_df$Variable))) +
  geom_hline(yintercept = 0, colour = gray(1/2), lty = 2) + 
  geom_linerange(aes(x = Variable, ymin = CI_low, ymax = CI_high),
                            lwd = 0.8, position = position_dodge(width = 1.2/2)) + 
  geom_linerange(aes(x = Variable, ymin = Q_25, ymax = Q_75),
                            lwd = 1.5, position = position_dodge(width = 1.2/2)) + 
  geom_pointrange(aes(x = Variable, y = Coefficient, ymin = Q_25,
                                 ymax = Q_75, fill = significant),
                             lwd = 1/2, shape = 21, position = position_dodge(width = 1.2/2)) +  
  scale_fill_manual(values = c("White", "black"))+
  facet_wrap(~ trait, scales = "free_x", ncol = 2) +
  coord_flip() + 
  scale_colour_viridis_d() +
  guides(fill = F, colour = F) +
  labs(colour = "Traits", x= "") +
  theme_mg_facets() +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.line = element_line(colour = "black"),
    legend.position = "bottom")

Plot_gamma

  plotfile <- paste0(model_wd, "/GammaPlots/Plot_gamma_", type_lab, "_", date, ".png")
  ggsave(plot = Plot_gamma, plotfile, height = 16, width = 20)
  
}

```



## Variance partitioning

Variance partitioning helps understand how much of the variation is explained by the environmental variables vs random effects (e.g., spatial structure, species-specific effects).

The order of the variables, if they are continuous, is 
1. intercept (this can be in any group)
2. first variable
3. second variable
ETC.

The formulas we used for running the models area: 
XFormula = ~ sex + BMI + Ntrapped + agri + imperv + tcd + swf + dist_road + 
    dist_path + cat_dens + pig_dens + pou_dens + precipitation_aug + 
    air_temp_aug

We grouped variables by different categories:  
- Mouse characteristics: sex, BMI,  
- Mouse density: Ntrapped,  
- Landcover: agri, imperv, tcd, swf_perc,  
- Connectivity: dist_road, dist_path,  
- Farm animals density: pou_dens, pig_dens, cat_dens,  
- Climatic: precipitation_aug, air_temp_aug.  

```{r variance partitioning}
# design matrix
head(m$X)

m$XFormula
# Total variance explained by hte model can be partition into the contributions of each fixed effect (or group) and random effect
# intercept can go in any group. Here in the first to simplify
# Groups: mouse_char, mouse_dens, environ, farm_animals, landcove, precipitation
VP <- computeVariancePartitioning(m, group = c(1, 1, 1, 2, 3, 3, 3, 3, 4, 4, 5, 5, 5, 6, 6),
                                  groupnames = c("mouse_characteristics", "mouse_dens", "landcover", "accessibility", "farm_animals", "climatic"))

# Save the Variance partitioning info
VP$vals
VP$R2T

saveRDS(VP, file = paste0(model_wd, "/jsdm_varpart_", date,".rds"))

```

```{r variance partitioning_2}
VP <- readRDS(paste0(model_wd, "/jsdm_varpart_", date,".rds"))

VP.table <- data.frame(VP$vals) %>%
  mutate(Factor = rownames(.)) %>%
  gather(., "ARO", "Variance_Explained", -Factor) %>%
  mutate(Factor = factor(Factor, 
                         levels = c("mouse_characteristics", "mouse_dens", "landcover","accessibility", "farm_animals",  "climatic", "Random: spatial_autocorrelation", "Random: farm_ID", "Random: year"),  
                         labels = c("Mouse characteristics","Mouse density", "Landcover", "Accessibility", "Farm animals density",  "Climatic", "Random: Spatial autocorrelation", "Random: Farm ID", "Random: Year")),
          ARO = gsub("\\.", ":", factor(ARO))) %>%
  left_join(., traits_summary) 

data_ordered_multiple <- VP.table %>%
  group_by(ARO) %>%
  mutate(access = Variance_Explained[Factor == "Accessibility"],
         farm = Variance_Explained[Factor == "Farm animals density"],
         landcover = Variance_Explained[Factor == "Landcover"],
         mousechar = Variance_Explained[Factor == "Mouse characteristics"],
         mousedens = Variance_Explained[Factor == "Mouse density"],
         climat = Variance_Explained[Factor == "Climatic"],
         sp_struc = Variance_Explained[Factor == "Random: Spatial autocorrelation"],
         farm = Variance_Explained[Factor == "Random: Farm ID"],
         year = Variance_Explained[Factor == "Random: Year"]) %>%
  ungroup() %>%
  arrange(
    desc(landcover),
    # desc(farm),
    # desc(landcover),
    # desc(mousechar),
    # desc(mousedens),
    year)

ordered_levels <- unique(data_ordered_multiple$ARO)
#ordered_levels


viridis_cols <- data.frame(colour_grp = setNames(viridis::viridis(length(unique(VP.table$Factor[!grepl("Random", VP.table$Factor)]))),
                         unique(VP.table$Factor[!grepl("Random", VP.table$Factor)])), Factor = unique(VP.table$Factor[!grepl("Random", VP.table$Factor)])) %>%
  mutate(Factor = factor(Factor, levels = c(
    "Mouse characteristics", "Mouse density", 
    "Landcover", "Accessibility",
    "Farm animals density",
    "Climatic", 
    "Random: Spatial autocorrelation", "Random: Farm ID", "Random: Year"
  )))

Mean_VP <- VP.table %>%
  group_by(Factor) %>%
  mutate(Factor = factor(Factor, levels = c(
    "Mouse characteristics", "Mouse density", 
    "Landcover", "Accessibility",
    "Farm animals density",
    "Climatic", 
    "Random: Spatial autocorrelation", "Random: Farm ID", "Random: Year"
  ))) %>%
  summarise(mean_var = round(mean(Variance_Explained), 2)) %>%
  arrange(desc(row_number())) %>% 
  ungroup() %>%
  left_join(., viridis_cols) %>%
  mutate(label = paste0(Factor, "\n", mean_var),
         ymin = c(0, head(cumsum(mean_var), -1)),
         ymax = cumsum(mean_var),
         texty = (ymin + ymax) / 2,
         colour_grp = ifelse(is.na(colour_grp), "grey20", colour_grp),
         alpha_group = ifelse(colour_grp=="grey20", 1:5, NA)) 


PLOT_piechart_avg_var <- ggplot(Mean_VP, aes(x = 2, y = mean_var, fill = Factor, alpha = Factor)) +
  geom_bar(stat = "identity", colour = "white") +
  geom_text_repel(aes(y = texty, label = mean_var, colour = colour_grp), size = 3, family = "Poppins", nudge_x = 1, 
            show_guide = FALSE) +
  scale_fill_manual(values = setNames(Mean_VP$colour_grp, Mean_VP$Factor)) +
  scale_colour_identity() +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(Mean_VP$alpha_group), to = c(1, 0.4), from = range(Mean_VP$alpha_group, na.rm = TRUE)), Mean_VP$Factor)) +
  coord_polar(theta = "y", start = 0) +
  labs(fill = "", alpha = "") +
  xlim(.5, 3) +
    guides(fill=guide_legend(ncol=3)) +
  theme_void() +
  theme(legend.position = "bottom",
        title = element_text(size = 10, family = "Poppins"),
        plot.margin = margin(0,0,0,0),
        legend.text = element_text(size = 9, family = "Poppins"))
  
PLOT_piechart_avg_var

ggsave(PLOT_piechart_avg_var, filename = paste0(model_wd, "/Plot_jsdm_VarPar_AVG", date, ".png"), 
       dpi = 600, height = 13.57, width = 13, units = "cm")



#####
VP.table <- VP.table %>%
  left_join(., Mean_VP)

PLOT_VarPart <- VP.table %>%
  ggplot(aes(x = factor(ARO, levels = ordered_levels), y = Variance_Explained, fill = Factor, alpha = Factor)) +
  scale_y_continuous(expand = c(0,0)) +
  geom_bar(stat = "identity", position = "stack", colour = "NA", width = 1) +
  scale_fill_manual(values = setNames(VP.table$colour_grp, VP.table$Factor)) +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(VP.table$alpha_group), to = c(0.3, 1), from = range(VP.table$alpha_group, na.rm = TRUE)),
  VP.table$Group_Label)) +
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "ARGs", y = "Variance Explained", fill = "", alpha = "") +
  theme_mg() +
  coord_flip() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 8),
        plot.text = element_text(size = 10)) 

PLOT_VarPart

# Save the faceted plot as a PNG
ggsave(PLOT_VarPart, filename = paste0(model_wd, "/Plot_jsdm_VarPar_", date, ".png"), 
       dpi = 600, height = 34, width = 20, units = "cm")

```

### Variance partitioning for mobile genes

```{r amr_varpart_mobile, echo=FALSE, fig.height=4, fig.width=6.3}

Mean_VP_mobile <- VP.table %>%
  filter(mobility == T, localization != "chromosome") %>%
  group_by(Factor) %>%
  summarise(mean_var = round(mean(Variance_Explained), 2)) %>%
  arrange(desc(row_number())) %>% 
  ungroup() %>%
  left_join(., viridis_cols) %>%
  mutate(label = paste0(Factor, "\n", mean_var),
         ymin = c(0, head(cumsum(mean_var), -1)),
         ymax = cumsum(mean_var),
         texty = (ymin + ymax) / 2,
         colour_grp = ifelse(is.na(colour_grp), "grey20", colour_grp),
         alpha_group = ifelse(colour_grp=="grey20", 1:3, NA)) %>%
  mutate(Factor = factor(Factor, levels = c(
    "Mouse characteristics", "Mouse density", 
    "Landcover", "Accessibility",
    "Farm animals density",
    "Climatic", 
    "Random: Spatial autocorrelation", "Random: Farm ID", "Random: Year"
  )))

PLOT_piechart_avg_var_mobile <- ggplot(data = Mean_VP_mobile, aes(x = 2, y = mean_var, fill = Factor, alpha = Factor)) +
  geom_bar(stat = "identity", colour = "white") +
  geom_text_repel(aes(y = texty, label = mean_var, colour = colour_grp), size = 3, family = "Poppins", nudge_x = 1, 
            show_guide = FALSE) +
  scale_fill_manual(values = setNames(Mean_VP_mobile$colour_grp, Mean_VP_mobile$Factor)) +
  scale_colour_identity() +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(Mean_VP_mobile$alpha_group), to = c(1, 0.4), from = range(Mean_VP_mobile$alpha_group, na.rm = TRUE)), Mean_VP_mobile$Factor)) +
  coord_polar(theta = "y", start = 0) +
  xlim(.5, 3) +
  labs(fill = "", alpha = "") +
  guides(fill=guide_legend(ncol=3)) +
  theme_void() +
  theme(legend.position = "bottom",
        title = element_text(size = 10, family = "Poppins"),
        plot.margin = margin(0,0,0,0),
        legend.text = element_text(size = 9, family = "Poppins"))
  
PLOT_piechart_avg_var_mobile


```

```{r amr_varpart_combined, echo=FALSE, fig.height=4, fig.width=6.3}

legend_var <- ggfun::get_legend(PLOT_piechart_avg_var_mobile)


PLOT_Avg_VarPart_tot_mob <- plot_grid(plot_grid(PLOT_piechart_avg_var + theme(legend.position = "none"), 
                                                PLOT_piechart_avg_var_mobile + theme(legend.position = "none"),
                                                align = "hv", 
                                                axis = "tblr",
                                                ncol = 2,
                                                labels = "auto"),
                                      legend_var,
                                      nrow = 2,
                                      rel_heights = c(1,0.4))

PLOT_Avg_VarPart_tot_mob

# Save the plot as a PNG

ggsave(PLOT_Avg_VarPart_tot_mob, filename = here("figures", paste0("Plot_jsdm_pa_varpart_tot_mob_", date, ".png")), height = 10, width = 16, units = "cm", dpi = 600)

# Save the plot as a SVG

ggsave(PLOT_Avg_VarPart_tot_mob, filename = here("figures", paste0("Plot_jsdm_pa_varpart_tot_mob_", date, ".svg")), height = 10, width = 16, units = "cm", dpi = 600)


```

### Variance partitioning for each traits

Here, we averaged the variance partitioning per traits variables.  

```{r variance partitioning in relation to traits}

Var_loc <- VP.table %>%
  group_by(localization, Factor) %>%
  summarize(Variance_Explained = mean(Variance_Explained)) %>% 
  ggplot(., aes(x = localization, y = Variance_Explained, fill = Factor, alpha = Factor)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = setNames(Mean_VP$colour_grp, Mean_VP$Factor)) +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(Mean_VP$alpha_group), to = c(0.3, 1), from = range(Mean_VP$alpha_group, na.rm = TRUE)), Mean_VP$Factor)) +
  labs(x = "Localization", y = "Variance Explained", fill = "Explaining Factor", alpha = "Explaining Factor",
       title = "") +
  theme_mg() +
    theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8),
        plot.text = element_text(size = 10),
        axis.title = element_text(size=10)) 

Var_drugclass <- VP.table %>%
  group_by(drug_class_grp, Factor) %>%
  summarize(Variance_Explained = mean(Variance_Explained)) %>% 
  ggplot(., aes(x = drug_class_grp, y = Variance_Explained, fill = Factor, alpha = Factor)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = setNames(Mean_VP$colour_grp, Mean_VP$Factor)) +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(Mean_VP$alpha_group), to = c(0.3, 1), from = range(Mean_VP$alpha_group, na.rm = TRUE)), Mean_VP$Factor)) +
  labs(x = "Drug Class", y = "Variance Explained", fill = "", alpha= "",
       title = "") +
  theme_mg() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.title = element_text(size=10))

Var_resmech <- VP.table %>%
  group_by(Resistance_Mechanism, Factor) %>%
  summarize(Variance_Explained = mean(Variance_Explained)) %>% 
  mutate(Resistance_Mechanism = str_replace_all(Resistance_Mechanism, c("antibiotic " = "",
                                                                        " to antibiotic" = ""))) %>%
  ggplot(., aes(x = Resistance_Mechanism, y = Variance_Explained, fill = Factor, alpha = Factor)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = setNames(Mean_VP$colour_grp, Mean_VP$Factor)) +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(Mean_VP$alpha_group), to = c(0.3, 1), from = range(Mean_VP$alpha_group, na.rm = TRUE)), Mean_VP$Factor)) +
  labs(x = "Resistance Mechanism", y = "Variance Explained", fill = "", alpha= "",
       title = "") +
  theme_mg() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.title = element_text(size=10))

legend_var <- ggfun::get_legend(Var_resmech)

Plot_VarPart_traits <- plot_grid(Var_loc + theme(legend.position = "none"), 
          Var_drugclass + theme(legend.position = "none"), 
          Var_resmech + theme(legend.position = "none"), 
          legend_var,
          align = "hv", 
          axis = "tblr",
          labels = c("a","b","c", ""))

# Save the plot as a PNG

ggsave(Plot_VarPart_traits, filename = here("figures", paste0("Plot_jsdm_pa_traits_varpart_", date, ".png")), height = 16, width = 16, units = "cm", dpi = 600)

# Save the plot as a SVG

ggsave(Plot_VarPart_traits, filename = here("figures", paste0("Plot_jsdm_pa_traits_varpart_", date, ".svg")), height = 16, width = 16, units = "cm", dpi = 600)


```



### Variance partitioning for individual variables

```{r variance partitioning}
# design matrix
head(m$X)

m$XFormula
# Total variance explained by hte model can be partition into the contributions of each fixed effect (or group) and random effect
# intercept can go in any group. Here in the first to simplify
# Groups: mouse_char, mouse_dens, environ, farm_animals, landcove, precipitation
VP_ind <- computeVariancePartitioning(m)

# Save the Variance partitioning info
VP_ind$vals
VP_ind$R2T

saveRDS(VP_ind, file = paste0(model_wd, "/jsdm_varpart_ind_", date,".rds"))


VP_ind <- readRDS(paste0(model_wd, "/jsdm_varpart_ind_", date,".rds"))

lvl <- c("sex","BMI","Ntrapped","agri","imperv","tcd","swf",
         "dist_road","dist_path",
         "cat_dens","pig_dens","pou_dens",
         "precipitation_aug","air_temp_aug",
         "Random: spatial_autocorrelation","Random: farm_ID","Random: year")

VP.table_ind <- data.frame(VP_ind$vals) %>%
  mutate(Factor = rownames(.)) %>%
  gather(., "ARO", "Variance_Explained", -Factor) %>%
  mutate(ARO = gsub("\\.", ":", factor(ARO))) %>%
  left_join(., traits_summary) %>%
    mutate(Factor = factor(Factor, levels = lvl)) 


viridis_cols_ind <- data.frame(colour_grp = setNames(viridis::turbo(length(unique(VP.table_ind$Factor[!grepl("Random", VP.table_ind$Factor)]))),
                         unique(VP.table_ind$Factor[!grepl("Random", VP.table_ind$Factor)])), Factor = unique(VP.table_ind$Factor[!grepl("Random", VP.table_ind$Factor)]))


  
Mean_VP_ind <- VP.table_ind %>%
  mutate(Factor = factor(Factor, levels = lvl)) %>%
  group_by(Factor) %>%
  summarise(mean_var = round(mean(Variance_Explained), 2)) %>%
  arrange(Factor) %>% 
  ungroup() %>%
  left_join(., viridis_cols_ind) %>%
  mutate(label = paste0(Factor, "\n", mean_var),
         ymin = c(0, head(cumsum(mean_var), -1)),
         ymax = cumsum(mean_var),
         texty = 1-(ymin + ymax) / 2,
         colour_grp = ifelse(is.na(colour_grp), "grey20", colour_grp),
         alpha_group = ifelse(colour_grp=="grey20", 1:5, NA)) 

PLOT_piechart_avg_var_ind <- Mean_VP_ind %>%
  ggplot(., aes(x = 2, y = mean_var, fill = Factor, alpha = alpha_group)) +
  geom_bar(stat = "identity", colour = "white") +
  geom_text_repel(aes(y = texty, label = label, colour = colour_grp), size = 4, family = "Poppins", nudge_x = 1.6) +
  scale_fill_manual(values = setNames(Mean_VP_ind$colour_grp, Mean_VP_ind$Factor)) +
  scale_colour_identity() +
  scale_alpha(range = c(1, 0.4), na.value = 1) +
  coord_polar(theta = "y", start = 0) +
  xlim(.5, 4) +
  theme_void() +
  theme(legend.position = "none",
        title = element_text(size = 10, family = "Poppins"),
        plot.margin = margin(0,0,0,0)) 
  
PLOT_piechart_avg_var_ind

ggsave(PLOT_piechart_avg_var_ind, filename = paste0(model_wd, "/Plot_jsdm_VarPar_AVG_IND", date, ".png"), 
       dpi = 600, height = 13.57, width = 13, units = "cm")

#####
VP.table_ind <- VP.table_ind %>%
  left_join(., Mean_VP_ind)

PLOT_VarPart_ind <- VP.table_ind %>%
  ggplot(aes(x = factor(ARO, levels = ordered_levels), y = Variance_Explained, fill = Factor, alpha = Factor)) +
  scale_y_continuous(expand = c(0,0)) +
  geom_bar(stat = "identity", position = "stack", colour = "NA", width = 1) +
  scale_fill_manual(values = setNames(VP.table_ind$colour_grp, VP.table_ind$Factor)) +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(VP.table_ind$alpha_group), to = c(0.3, 1), from = range(VP.table_ind$alpha_group, na.rm = TRUE)),
  VP.table_ind$Group_Label)) +
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "ARGs", y = "Variance Explained", fill = "", alpha = "") +
  theme_mg() +
  coord_flip() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 8),
        plot.text = element_text(size = 10)) 

PLOT_VarPart_ind

# Save the faceted plot as a PNG
ggsave(PLOT_VarPart_ind, filename = paste0(model_wd, "/Plot_jsdm_VarPar_individual_var_", date, ".png"), 
       dpi = 600, height = 15, width = 34, units = "cm")
```



## Results for clinically relevant genes 

```{r clinically relevant genes extraction}

arg_clinrel <- read.csv("./data/raw/arg_clinically_relevant.csv")
head(arg_clinrel)

# identify ARGs that are clinically relevant in our data

arg_clinrel_group <- arg_card_short %>%
  mutate(key = 1) %>%
  inner_join(arg_clinrel %>% mutate(key = 1), by = "key") %>%
  filter(str_detect(ARG_CARD_Short, regex(Gen_groups, ignore_case = TRUE))) %>%
  dplyr::select(ARO, Gen_groups) %>%
  distinct() %>%
  left_join(., traits_summary, by = c("ARO")) %>%
  filter(!Gen_groups == c("arr"))

arg_clinrel_list <- arg_clinrel_group %>%
  pull(ARO) 

arg_clinrel_list

# 72 ARO selected as clinically relevant

```

### Estimates

```{r filtering results}

postBeta <- readRDS(paste0(model_wd, "/jsdm_postBeta_", date, ".rds"))

#Get species and covariate names from model
argNames <- gsub( "\\.", ":", colnames(m$Y))
covNames <- m$covNames

#pull the data from postGamma for plotting
mBeta <- postBeta$mean
BetaP <- postBeta$support

Beta_association_mean <- data.frame(mBeta)
colnames(Beta_association_mean) <- argNames
rownames(Beta_association_mean) <- covNames
#Beta_association_values[Beta_association_values == 0] <- NA

Beta_association_support <- data.frame(BetaP)
colnames(Beta_association_support) <- argNames
rownames(Beta_association_support) <- covNames

Beta_association_values_df_clinrel <- Beta_association_mean %>%
  mutate(covNames = rownames(Beta_association_mean)) %>%
  gather(ARO, value, -covNames) %>%
  left_join(., Beta_association_support %>%
  mutate(covNames = rownames(Beta_association_support)) %>%
  gather(ARO, support_lvl, -covNames)) %>%
  filter(ARO %in% arg_clinrel_list) %>%
  mutate(covNames = factor(covNames, levels = c("(Intercept)", "sexM", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precipitation_aug", "air_temp_aug"),
                           labels = c("(Intercept)", "sex-M", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precip", "temp"))) %>%
  left_join(., arg_clinrel_group, by = "ARO") %>%
  group_by(Gen_groups, Resistance_Mechanism, covNames) %>%
  summarise(value = mean(value),
            support_lvl = mean(support_lvl)) %>%
  mutate(sign = ifelse(support_lvl > 0.95 | support_lvl < (1 - 0.95), "*", 
                       ifelse(support_lvl > 0.75 | support_lvl < (1 - 0.75), "·", NA)))
  


PLOT_Est_clinrel <- Beta_association_values_df_clinrel %>%
  #full_join(N_arg) %>%
  mutate(covNames = factor(covNames, levels = c("(Intercept)","sex-M", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precip", "temp", "# ARG", "# mice","# farm"))) %>%
  mutate(group = case_when(covNames %in% c("sex-M", "BMI") ~ "Mouse\ncharacteristics",
                           covNames == c("Ntrapped") ~ "Mouse\ndensity",
                           covNames %in% c("agri", "imperv", "tcd", "swf") ~ "Landcover",
                           covNames %in% c("dist_road", "dist_path") ~ "Accessibility",
                           covNames %in% c("cat_dens", "pig_dens", "pou_dens") ~ "Farm animal\ndensity",
                           covNames %in% c("precip", "temp") ~ "Climatic",
                           covNames %in% c("# ARG", "# mice","# farm") ~ "Sample\nsize"),
         group = factor(group, levels = c("Mouse\ncharacteristics","Mouse\ndensity","Landcover","Accessibility","Farm animal\ndensity","Climatic","Sample\nsize"))) %>%
  filter(covNames != "(Intercept)") %>%
  ggplot(., aes(y = Gen_groups, x = covNames, fill = value)) +
  geom_tile(colour = "black") +
  geom_text(aes(label = sign), size = 5, colour = "black") +
  #geom_text(aes(label = sample), size = 2.5, family = "Poppins") +
  #scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-4.5, 10.5)) +
  scale_fill_gradientn(colors = c("blue4", "blue3","white", "red3","red4"),
                       values = scales::rescale(c(-5,-2, 0, 5, 10)),
                       breaks = c(-5,0,5,10),
                       limits = c(-5,10.5),
                       na.value="grey99") +
  scale_x_discrete(position = "top", guide = guide_axis(angle = 90)) +
#  facet_col(type ~ group, scales = "free_x",  space = "free", strip.position = c("top", "right")) +
  facet_grid(Resistance_Mechanism ~ group, scales = "free", space = "free", switch = "y") +
  labs(fill = "Coefficient estimate") +
  theme_mg_facets() +
  theme(
    text = element_text(size = 8),
    axis.text.y = element_text(hjust = 1, vjust = 0.4, size = 8),
    axis.text.x = element_text(size = 8),
    axis.title = element_blank(),
    axis.line = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    strip.placement = "outside",
    strip.background = element_rect(fill = "grey90", colour = "grey"),
    strip.text.x = element_text(angle = 90, size = 8),
    strip.text.y.left = element_text(angle = 0, size = 8),
    panel.spacing = unit(0, "cm"))

PLOT_Est_clinrel

# Save the plot as a PNG

ggsave(PLOT_Est_clinrel, filename = here("figures", paste0("Plot_jsdm_pa_clinrel_est_", date, ".png")), height = 20, width = 16, units = "cm", dpi = 600)

# Save the plot as a SVG

ggsave(PLOT_Est_clinrel, filename = here("figures", paste0("Plot_jsdm_pa_clinrel_est_", date, ".svg")), height = 20, width = 16, units = "cm", dpi = 600)

######### Alternative plot

PLOT_Est_clinrel <- Beta_association_values_df_clinrel %>%
  #full_join(N_arg) %>%
  mutate(covNames = factor(covNames, levels = c("(Intercept)","sex-M", "BMI", "Ntrapped", "agri", "imperv", "tcd", "swf", "dist_road", "dist_path", "cat_dens", "pig_dens", "pou_dens", "precip", "temp", "# ARG", "# mice","# farm"))) %>%
  mutate(group = case_when(covNames %in% c("sex-M", "BMI") ~ "Mouse\ncharacteristics",
                           covNames == c("Ntrapped") ~ "Mouse\ndensity",
                           covNames %in% c("agri", "imperv", "tcd", "swf") ~ "Landcover",
                           covNames %in% c("dist_road", "dist_path") ~ "Accessibility",
                           covNames %in% c("cat_dens", "pig_dens", "pou_dens") ~ "Farm animal\ndensity",
                           covNames %in% c("precip", "temp") ~ "Climatic",
                           covNames %in% c("# ARG", "# mice","# farm") ~ "Sample\nsize"),
         group = factor(group, levels = c("Mouse\ncharacteristics","Mouse\ndensity","Landcover","Accessibility","Farm animal\ndensity","Climatic","Sample\nsize"))) %>%
  filter(covNames != "(Intercept)") %>%
  ggplot(., aes(y = Resistance_Mechanism, x = covNames, fill = value)) +
  geom_tile(colour = "black") +
  geom_text(aes(label = sign), size = 5, colour = "black") +
  #geom_text(aes(label = sample), size = 2.5, family = "Poppins") +
  #scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,limits = c(-4.5, 10.5)) +
  scale_fill_gradientn(colors = c("blue4", "blue3","white", "red3","red4"),
                       values = scales::rescale(c(-5,-2, 0, 5, 10)),
                       breaks = c(-5,0,5,10),
                       limits = c(-5,10.5),
                       na.value="grey99") +
  scale_x_discrete(position = "top", guide = guide_axis(angle = 90)) +
#  facet_col(type ~ group, scales = "free_x",  space = "free", strip.position = c("top", "right")) +
  facet_grid(Gen_groups ~ group, scales = "free", space = "free", switch = "y") +
  labs(fill = "Coefficient estimate") +
  theme_mg_facets() +
  theme(
    text = element_text(size = 8),
    axis.text.y = element_text(hjust = 1, vjust = 0.4, size = 8),
    axis.text.x = element_text(size = 8),
    axis.title = element_blank(),
    axis.line = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8),
    strip.placement = "outside",
    strip.background = element_rect(fill = "grey90", colour = "grey"),
    strip.text.x = element_text(angle = 90, size = 8),
    strip.text.y.left = element_text(angle = 0, size = 8),
    panel.spacing = unit(0, "cm"))

PLOT_Est_clinrel

# Save the plot as a PNG

ggsave(PLOT_Est_clinrel, filename = here("figures", paste0("Plot_jsdm_pa_clinrel_est_alt_", date, ".png")), height = 20, width = 16, units = "cm", dpi = 600)

# Save the plot as a SVG

ggsave(PLOT_Est_clinrel, filename = here("figures", paste0("Plot_jsdm_pa_clinrel_est_alt_", date, ".svg")), height = 20, width = 16, units = "cm", dpi = 600)




```


### Variance partitionning

```{r variance partitioning}

VP <- readRDS(paste0(model_wd, "/jsdm_varpart_", date,".rds"))

VP.table_clinrel <- data.frame(VP$vals) %>%
  mutate(Factor = rownames(.)) %>%
  gather(., "ARO", "Variance_Explained", -Factor) %>%
  mutate(Factor = factor(Factor, 
                         levels = c("mouse_characteristics", "mouse_dens", "landcover","accessibility", "farm_animals",  "climatic", "Random: spatial_autocorrelation", "Random: farm_ID", "Random: year"),  
                         labels = c("Mouse characteristics","Mouse density", "Landcover", "Accessibility", "Farm animals density",  "Climatic", "Random: Spatial autocorrelation", "Random: Farm ID", "Random: Year")),
          ARO = gsub("\\.", ":", factor(ARO))) %>%
  left_join(., arg_clinrel_group, by = join_by(ARO == arg)) %>%
  filter(ARO %in% arg_clinrel_list) %>%
  distinct()


viridis_cols_clinrel <- data.frame(colour_grp = setNames(viridis::viridis(length(unique(VP.table_clinrel$Factor[!grepl("Random", VP.table_clinrel$Factor)]))),
                         unique(VP.table_clinrel$Factor[!grepl("Random", VP.table_clinrel$Factor)])), Factor = unique(VP.table_clinrel$Factor[!grepl("Random", VP.table_clinrel$Factor)])) %>%
  mutate(Factor = factor(Factor, levels = c(
    "Mouse characteristics", "Mouse density", 
    "Landcover", "Accessibility",
    "Farm animals density",
    "Climatic", 
    "Random: Spatial autocorrelation", "Random: Farm ID", "Random: Year"
  )))


Mean_VP_clinrel <- VP.table_clinrel %>%
  mutate(Factor = factor(Factor)) %>%
  group_by(Factor) %>%
  summarise(mean_var = round(mean(Variance_Explained), 2)) %>%
  arrange(Factor) %>% 
  ungroup() %>%
  left_join(., viridis_cols_clinrel) %>%
  mutate(label = paste0(Factor, "\n", mean_var),
         ymin = c(0, head(cumsum(mean_var), -1)),
         ymax = cumsum(mean_var),
         texty = 1-(ymin + ymax) / 2,
         colour_grp = ifelse(is.na(colour_grp), "grey20", colour_grp),
         alpha_group = ifelse(colour_grp=="grey20", 1:3, NA)) 

PLOT_piechart_avg_var_clinrel <- Mean_VP_clinrel %>%
  ggplot(., aes(x = 2, y = mean_var, fill = Factor, alpha = alpha_group)) +
  geom_bar(stat = "identity", colour = "white") +
  geom_text_repel(aes(y = texty, label = label, colour = colour_grp), size = 4, family = "Poppins", nudge_x = 1.6) +
  scale_fill_manual(values = setNames(Mean_VP_clinrel$colour_grp, Mean_VP_clinrel$Factor)) +
  scale_colour_identity() +
  scale_alpha(range = c(1, 0.4), na.value = 1) +
  coord_polar(theta = "y", start = 0) +
  xlim(.5, 4) +
  theme_void() +
  theme(legend.position = "none",
        title = element_text(size = 10, family = "Poppins"),
        plot.margin = margin(0,0,0,0)) 
  
PLOT_piechart_avg_var_clinrel

ggsave(PLOT_piechart_avg_var_clinrel, filename = paste0(model_wd, "/Plot_jsdm_VarPar_AVG_CLINREL", date, ".png"), 
       dpi = 600, height = 13.57, width = 13, units = "cm")

# Save the faceted plot as a PNG
ggsave(PLOT_piechart_avg_var_clinrel, filename = here("figures", paste0("Plot_jsdm_VarPar_tot_clinrel_", date, ".png")), dpi = 600, height = 13.57, width = 13, units = "cm")

# Save the faceted plot as a SVG
ggsave(PLOT_piechart_avg_var_clinrel, filename = here("figures", paste0("Plot_jsdm_VarPar_tot_clinrel_", date, ".svg")), dpi = 600, height = 13.57, width = 13, units = "cm")




#####
VP.table_clinrel <- VP.table_clinrel %>%
  group_by(Factor, Gen_groups) %>%
  summarise(Variance_Explained = mean(Variance_Explained)) %>%
  left_join(., Mean_VP_clinrel)


PLOT_VarPart_clinrel <- VP.table_clinrel %>%
  ggplot(aes(x = factor(Gen_groups), y = Variance_Explained, fill = Factor, alpha = Factor)) +
  scale_y_continuous(expand = c(0,0)) +
  geom_bar(stat = "identity", position = "stack", colour = "NA", width = 1) +
  scale_fill_manual(values = setNames(VP.table_clinrel$colour_grp, VP.table_clinrel$Factor)) +
  scale_alpha_manual(values = setNames(scales::rescale(as.numeric(VP.table_clinrel$alpha_group), to = c(0.3, 1), from = range(VP.table_clinrel$alpha_group, na.rm = TRUE)), VP.table_clinrel$Factor)) +
  guides(fill=guide_legend(ncol=2)) +
  labs(x = "ARG groups", y = "Variance Explained", fill = "", alpha = "") +
  theme_mg() +
  coord_flip() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 8),
        plot.text = element_text(size = 10)) 

PLOT_VarPart_clinrel

# Save the faceted plot as a PNG
ggsave(PLOT_VarPart_clinrel, filename = here("figures", paste0("Plot_jsdm_VarPar_clinrel_", date, ".png")), dpi = 600, height = 15, width = 16, units = "cm")

# Save the faceted plot as a SVG
ggsave(PLOT_VarPart_clinrel, filename = here("figures", paste0("Plot_jsdm_VarPar_clinrel_", date, ".svg")), dpi = 600, height = 15, width = 16, units = "cm")

```






## Species associations

We do it through species associations in residual random variance.   
One of the strengths of HMSC is its ability to model species associations. You can analyze the species covariance matrix to visualize these associations.  

```{r sp associations}
OmegaCor <- computeAssociations(m)
class(OmegaCor)
saveRDS(OmegaCor, file=paste0(model_wd, "/jsdm_OmegaCor_", date, ".rds"))

# OmegaCor <- readRDS(paste0(model_wd, "/jsdm_OmegaCor_", date, ".rds"))

OmegaCor[[1]]$mean
OmegaCor[[1]]$support

# Default plot in Hmsc package
supportLevel <- 0.95

association_values <- data.frame(((OmegaCor[[1]]$support > supportLevel)
                       + (OmegaCor[[1]]$support < (1 - supportLevel)) > 0) * OmegaCor[[1]]$mean)

PLOT_Cor_ARG_resmech_95 <- ggcorrplot(association_values, method = "square") +
  labs(colour = "Correlation") +
  theme_mg_cor() 

PLOT_Cor_ARG_resmech_95

ggsave(paste0(model_wd, "/Plot_jsdm_default_omegaplot95_", date,".png"), PLOT_Cor_ARG_resmech_95, dpi = 600)

##
supportLevel <- 0.75

association_values <- data.frame(((OmegaCor[[1]]$support > supportLevel)
                       + (OmegaCor[[1]]$support < (1 - supportLevel)) > 0) * OmegaCor[[1]]$mean)

PLOT_Cor_ARG_resmech_75 <- ggcorrplot(association_values, method = "square") +
  labs(colour = "Correlation") +
  theme_mg_cor() 

PLOT_Cor_ARG_resmech_75

ggsave(paste0(model_wd, "/Plot_jsdm_default_omegaplot75_", date,".png"), PLOT_Cor_ARG_resmech_75, dpi = 600)

```

```{r save associations}
assoc.mean <- melt(OmegaCor[[1]]$mean)
assoc.support <- melt(OmegaCor[[1]]$support)

nrow(assoc.mean); nrow(assoc.support)

associations <- cbind.data.frame(assoc.mean, support = assoc.support$value)
colnames(associations) <- c("species1", "species2", "mean", "support")

associations <- associations %>%
  mutate(species1 = gsub(".", " ", species1, fixed =TRUE),
         species2 = gsub(".", " ", species2, fixed =TRUE))

write.csv(associations, paste0(model_wd, "/jsdm_species_associations_", date, ".csv"), row.names = FALSE)
```


### Circular plot 

Plot the species associations categorizing the taxa by some grouping characteristic
We use a Hierarchical edge bundling plot

To create a bundling graph we need a hierarchy that is going to be used for the dendrogram. 
We add higher categories to our species, grouping them by env response, that we will use later for plotting

We also need to create the edges: data frame giving the hierarchical structure of the species,
grouped by species group


```{r load data}
# Taxa 
response <- as.data.frame(m$Y)

## create random grouping as example
# taxa
my_groups <- data.frame(species = unique(associations$species1))
# grouping in three groups, as an example, no need 
my_groups$group <- factor(rep("1", 340))
# n sites it was detected
nsites <- data.frame(species = gsub(".", " ", colnames(response), fixed = TRUE), 
                     nsites = colSums(response))
my_groups <- my_groups %>% 
  left_join(nsites, by = "species")

my_groups

## Get sp names for the plots
associations$sp1 <- as.character(associations$species1)
associations$sp2 <- as.character(associations$species2)
  
head(associations, 20)  

# Select association that are significant in the 95% CI
assoc_corr95 <- associations %>%
  filter(support < 0.025 | support > 0.975) %>%
  filter(species1 != species2) %>%
  dplyr::select(from = sp1, to = sp2, mean = mean, support = support) %>%
  mutate(correl = ifelse(mean < 0, "neg", "pos"))

nrow(assoc_corr95)
head(assoc_corr95)
```

```{r prepare hierarchical edge bundling plot}
# groups
d1 <- data.frame(from = "origin", to = levels(my_groups$group))

# Then we assign the species to the groups
d2 <- data.frame(from = my_groups$group, to = my_groups$species)
edges <- rbind(d1, d2)
nrow(edges)

## prepare nodes
#We use the number of sites that the species was detected for the size of the dots
vertices <- data.frame(name = unique(c(as.character(edges$from), as.character(edges$to))))
nrow(vertices)

vertices$size <- c(rep(NA, 2), my_groups$nsites)
vertices$group = edges$from[match( vertices$name, edges$to )]
head(vertices, 10)

# Reorder the data following the order we are interested in for representation:
# so all the species in the same group appear together
edges.ord <- edges[order(edges$from),]
vertices.ord <- vertices[order(vertices$group,vertices$name),]
head(vertices.ord, 10)

## calculate the ANGLE of the labels for plotting
vertices.ord$id=NA
myleaves <- which(is.na(match(vertices.ord$name, edges.ord$from))) # Select the rows with the final species, not the groups
nleaves <- length(myleaves) # This should be the number of species
vertices.ord$id[myleaves] <- seq(1:nleaves)

# First angle to calculate horizontal adjustment
vertices.ord$angle <- 90 - 360*vertices.ord$id/nleaves

# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
vertices.ord$hjust <- ifelse(vertices.ord$angle < -90, 1, 0)

# flip angle BY to make them readable
vertices.ord$angle <- ifelse(vertices.ord$angle < -90, vertices.ord$angle + 180, vertices.ord$angle)

head(vertices.ord, 10)

vertices.ord <- vertices.ord %>%
  mutate(name = gsub(".", " ", name, fixed =TRUE))

```

```{r create graph object and plot}
# give data graph format
mygraph <- graph_from_data_frame(d = edges.ord, vertices = vertices.ord)

# The connection object must refer to the ids of the leaves:
from <- match(assoc_corr95$from, vertices.ord$name)
to <- match(assoc_corr95$to, vertices.ord$name)

# we transform the correlations into positive or negative to better plot
connections <- assoc_corr95 %>%
  mutate(correlation = ifelse(mean > 0, "positive", "negative"))
connections

my.plot <- ggraph(mygraph, layout = 'dendrogram', circular = TRUE) +
  geom_conn_bundle(data = get_con(from = from, to = to, values = connections$correlation), 
                   alpha = 0.3, width = 1.2, tension = 0.5, aes(colour = factor(values)), 
                   show.legend = TRUE) +
  geom_node_text(aes(x = x*1.2, y = y*1.2, filter = leaf, label = name, angle = 0, 
                     hjust = hjust), size = 3.5, alpha = 0.7, family = "Poppins") +
  geom_node_point(aes(filter = leaf, x = x*1.07, y = y*1.07, #colour = group, 
                      size = size, alpha=0.5)) +  
  #expand_limits(x = c(-3, 3), y = c(-2, 2)) +
  expand_limits(x = c(-1.5, 1.5)) +
  # Set the color palette for the edges
  scale_edge_colour_manual(values = c("positive" = viridis_pal()(2)[1], 
                                      "negative" = viridis_pal()(2)[2]), 
                           name = "Association:") +
  scale_edge_linetype("pointed") +
  scale_size_continuous(range = c(1,8),
                        breaks = c(100, 250, 500, 750),
                        name = "Number of samples:") +
  guides(size = guide_legend(order = 2),
         colour = "none",
         alpha = "none") +
  # change theme  
  theme_mg_cor_cir() 

my.plot

ggsave(paste0(model_wd, "/Plot_species_associations_circular_", date, ".png"), plot = my.plot,
       dpi = 600, width = 15.79, height = 13.33, units = "cm")

```

