---
title: "CompareLivestockData"
author: "Aimara Planillo and Moritz Wenzler-Meya"
date: "2026-03-24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Compare livestock data from different years

```{r}
source("scripts/source_packages.R")
```


### Load data

```{r data}
data <- read.csv("data/processed/Table_ARG_mice_envcov_100m_202507.csv")

# livestock <- st_read("data/raw/geo/thuenen_livestock/livestock_germany_1999-2020_polygon_04326.gpkg")
# ## get data to use
# livestock_filtered <- livestock %>% 
#   dplyr::select(CATO_2016, CATO_2020,
#                 PITO_2016, PITO_2020, 
#                 POTO_2016, POTO_2020, 
#                 geom)
# 
# st_crs(livestock_filtered)

# st_write(livestock_filtered, "./data/processed/geo/livestock_2016vs2020_31467.gpkg")

livestock_filtered <- st_read("./data/processed/geo/livestock_2016vs2020_31467.gpkg")
```

## get mice points
```{r}
head(data)

data_sf <- st_as_sf(data, coords = c("lon", "lat"), crs = 4326) %>% 
  st_transform(st_crs(livestock))
```

Livestock units in 1000 GVE; 
**This value represents the amount of livestock measured in "1000 livestock units" (1000000 units for poultry)**
As it is standardized, can be directly comparable between regions 
CATO: Cattle; the sum of all cattle related subclasses
PITO: Pigs; the sum of all pig related subclasses
POTO: Poultry; the sum of all poultry related subclasses


## extract farm animals in farms
```{r}
data_sf_ext <- st_intersection(data_sf, livestock_filtered)

## get info at the farm level
data_sf_ext_farm <- data_sf_ext %>% 
  group_by(farm_ID) %>% 
  summarise(across(CATO_2016:POTO_2020, mean))
```


## comparison pigs: PITO
```{r}
cor_pito <- cor(x = data_sf_ext_farm$PITO_2016, 
                y = data_sf_ext_farm$PITO_2020, 
                use = "pairwise.complete.obs",
                method = "pearson")
max_limit <- max(data_sf_ext_farm$PITO_2016, data_sf_ext_farm$PITO_2020, na.rm = TRUE)

ggplot(data_sf_ext_farm, aes(x=PITO_2016, y=PITO_2020)) +
  geom_point(col = "steelblue", alpha = 0.50, size = 4, pch = 16) +
  geom_abline(slope = cor_pito, intercept = 0, col = "grey80", lwd = 2, lty = 4) +
  # geom_smooth(method=lm , color="red", se=FALSE) + 
  coord_equal(xlim = c(0, max_limit), ylim = c(0, max_limit)) +
  labs(title = "Comparison pigs relative abundance 2016 vs 2020",
       subtitle = paste0("Pearson Correlation = ", round(cor_pito, digits = 4)),
       x = "Pigs - relative abundance in 2020",
       y = "Pigs - relative abundance in 2016") +
  theme_bw() +
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12))

ggsave("figures/comparison_2016_2020/comparison_pigs_2016_vs_2020.jpg", dpi = 300, width = 7, height = 7)
```

comparison cattle: CATO
```{r}
cor_cato <- cor(x = data_sf_ext_farm$CATO_2016, 
                y = data_sf_ext_farm$CATO_2020, 
                use = "pairwise.complete.obs",
                method = "pearson")
max_limit <- max(data_sf_ext_farm$CATO_2016, data_sf_ext_farm$CATO_2020, na.rm = TRUE)

ggplot(data_sf_ext_farm, aes(x=CATO_2016, y=CATO_2020)) +
  geom_point(col = "steelblue", alpha = 0.50, size = 4, pch = 16) +
  geom_abline(slope = cor_cato, intercept = 0, col = "grey80", lwd = 2, lty = 4) +
  # geom_smooth(method=lm , color="red", se=FALSE) + 
  coord_equal(xlim = c(0, max_limit), ylim = c(0, max_limit)) +
  labs(title = "Comparison cattle relative abundance 2016 vs 2020",
       subtitle = paste0("Pearson Correlation = ", round(cor_cato, digits = 4)),
       x = "Cattle - relative abundance in 2020",
       y = "Cattle - relative abundance in 2016") +
  theme_bw() +
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12))

ggsave("figures/comparison_2016_2020/comparison_cattle_2016_vs_2020.jpg", dpi = 300, width = 7, height = 7)
```

comparison poultry: POTO

```{r}
cor_poto <- cor(x = data_sf_ext_farm$POTO_2016, 
                y = data_sf_ext_farm$POTO_2020, 
                use = "pairwise.complete.obs",
                method = "pearson")
max_limit <- max(data_sf_ext_farm$POTO_2016, data_sf_ext_farm$POTO_2020, na.rm = TRUE)

ggplot(data_sf_ext_farm, aes(x=POTO_2016, y=POTO_2020)) +
  geom_point(col = "steelblue", alpha = 0.50, size = 4, pch = 16) +
  geom_abline(slope = cor_poto, intercept = 0, col = "grey80", lwd = 2, lty = 4) +
  # geom_smooth(method=lm , color="red", se=FALSE) + 
  coord_equal(xlim = c(0, max_limit), ylim = c(0, max_limit)) +
  labs(title = "Comparison poultry relative abundance 2016 vs 2020",
       subtitle = paste0("Pearson Correlation = ", round(cor_poto, digits = 4)),
       x = "Poultry - relative abundance in 2020",
       y = "Poultry - relative abundance in 2016") +
  theme_bw() +
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12))

ggsave("figures/comparison_2016_2020/comparison_poultry_2016_vs_2020.jpg", dpi = 300, width = 7, height = 7)
```


```{r}
test <- st_read("./data/processed/geo/livestock_germany_2020_vector_3035.gpkg")



cbind.data.frame(livestock$PITO_2020, test$PITO)
```


